### 2015-04-28 completeNull/maxt.tex
added directories for subprojects

\documentclass{article}
\usepackage[left=1.25in,top=1.25in,right=1.25in,bottom=1.25in,head=1.25in]{geometry}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{verbatim,float,url,enumerate}
\usepackage{graphicx,subfigure,epsfig,psfrag}
\usepackage{bm,dsfont,color,appendix}
\usepackage{natbib}
\usepackage{latexsym,graphicx}
%\usepackage{showkeys}

\newtheorem{algorithm}{Algorithm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{corollary}{Corollary} 
\newtheorem{proposition}{Proposition}

\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\maximize}{\mathop{\mathrm{maximize}}}
\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}

\newcommand{\mm}{\mathcal{M}}
\newcommand{\pp}{\mathcal{P}}

\newcommand{\inc}{{\rm inc}}
\newcommand{\comp}{{\rm comp}}
\newcommand{\FM}{ {\rm FM}}
% \def\baselinestretch{1.7931}  
\pdfminorversion=4

\def\real{\mathbb{R}}
\def\epi{\mathbf{epi}}
\def\E{\mathbb{E}}
\def\P{\mathbb{P}}
\def\Var{\mathrm{Var}}
\def\Cov{\mathrm{Cov}}
\def\half{\frac{1}{2}}
\def\th{\mathrm{th}}
\def\tr{\mathrm{tr}}
\def\df{\mathrm{df}}
\def\dim{\mathrm{dim}}
\def\col{\mathrm{col}}
\def\row{\mathrm{row}}
\def\nul{\mathrm{null}}
\def\rank{\mathrm{rank}}
\def\nuli{\mathrm{nullity}}
\def\sign{{\mathrm{sign}}}
\def\supp{\mathrm{supp}}
\def\diag{\mathrm{diag}}
\def\aff{\mathrm{aff}}
\def\hy{\hat{y}}
\def\by{{\bf y}}
\def \sI {{\cal I}}
\def\bfeta{{\bf \eta}}
\def\bmu{{\bf \mu}}
\def\ty{{\tilde y}}
\newcommand{\Nu}{{\cal V}}
\def\bI {{\bf I}}
\def\by{{\bf y}}
\def\bSigma {{\bf\Sigma}}
\def\ty{\tilde{y}}
\def\hbeta{\hat{\beta}}
\def\tbeta{\tilde{\beta}}
\def\hu{\hat{u}}
\def\hv{\hat{v}}
\def\lone{1}
\def\ltwo{2}
\def\linf{\infty}
\def\lzero{0}
\def\T{^T}
\def\R{\mathbb{R}}
\def\cA{\mathcal{A}}
\def\cB{\mathcal{B}}
\def\cD{\mathcal{D}}
\def\cE{\mathcal{E}}
\def\cF{\mathcal{F}}
\def\cG{\mathcal{G}}
\def\cM{\mathcal{M}}
\def\cN{\mathcal{N}}
\def\cO{\mathcal{O}}
\def\cP{\mathcal{P}}
\def\cS{\mathcal{S}}
\def\cT{\mathcal{T}}
\def\cZ{\mathcal{Z}}
\def\Exp{\mathrm{Exp}}
\def\LARS{\mathrm{LARS}}
\def\RSS{\mathrm{RSS}}
\def\spa{\mathrm{span}}
\def\cl{\mathrm{cl}}
\def\relint{\mathrm{relint}}
\def\lin{\mathrm{lin}}
\def\codim{\mathrm{codim}}
\def\F{\mathbb{F}}

\usepackage[T1]{fontenc}
\def\qu{\textquotedbl}

\newcommand{\cd}{\overset{d}{\rightarrow}}
\newcommand{\sphere}[1]{O^{#1-1}}
\newcommand{\hittime}[2]{\gamma(#1,#2)}
\def\strong{S}
\def\lips{\mathcal{L}}
\def\dimens{\text{dim}}
\def\cone{\text{Cone}}
\def\critval{\gamma}
\def\tube{\mathrm{Tube}}
\def\V{\mathcal{V}}
\def\K{\mathcal{K}}
\def\numsignals{L}
\def\numevents{E}
\def\weight{\omega}
\def\winv{\theta}
\def\Pp{\mathbb{P}}
\def\Gfirst{G}
\def\Gsecond{{\bar{G}}}
\def\modelinfo{{\cal Z}}
\def\cS {\cal S}
\def\Prob{{\rm Pr}}

\title{	Bootstrap and analytic conditional tests for  forward stepwise regression}
%\author{Ryan Tibshirani$^1$, Robert J. Tibshirani$^2$, Alessandro Rinaldo$^1$ \and Larry Wasserman$^1$  in some order}
\author{Tibs times two}

%\and Robert Tibshirani$^1$}
%\date{\normalsize $^1$Stanford University,
%$^2$Simon Fraser University, 
%$^3$Carnegie Mellon University}

\begin{document}
\maketitle

\begin{abstract}
We study the standard ``max-t'' test for sequential testing in forward stepwise regression.
We provide computational and analytic approaches for carrying out this test,  that yield type- I error
control  conditional on the selection inherent in the procedure.
We compare the power of this procedure to the max-t test of Buja et al. and the recently proposed ``spacing'' test of \citet{TLTT2014}.
We propose two versions of the procedure: one that uses a  Gaussian error model and the other a  nonparametric bootstrap approach.

Keywords: Forward stepwise regression, multiple testing

\end{abstract}

\section{Introduction}
\label{sec:intro}
We consider observations $y \in \real^n$ drawn from a Gaussian model  
\begin{equation}
y = \theta + \epsilon, \;\;\; \epsilon \sim N(0,\sigma^2 I),
\label{eq:mod}
\end{equation}
Given a fixed matrix $X \in \real^{n\times p}$ of predictor variables,
we  derive new inferential tools for forward stepwise regression, that properly adjust
for the selection inherent to the procedure.

Assume that $X$ has standardized columns and $y$ has mean zero.
Suppose that we run forward stepwise regression for $k-1$ steps, yielding a model $M$ having  least squares coefficients
$\hat\beta$ and residuals $r=y-X\hat\beta$. Let $\tilde X=P_M^\perp X$, the residuals of each column of $X$ after projection onto $M$.
The  standard ``Max'-t" test for entering the next predictor defines
$$ T_j=\frac{\tilde x_j^Tr}{||\tilde x_j||};\; j\notin M$$
and chooses $\hat j ={\rm argmax}_{j\notin M}  |T_j|$.
In this paper we establish some  optimal properties of this sequential test, conditional on the selection, and provide analytic formulae for its implementation.
We call this procedure the {\em Conditional max-t} test.

We wish to stress that this test, of  course, is far from new. The max-t test is the standard test for forward stepwise regression.
Typically it is used without any attempt to adjust for the  effects of selection, that is, to test significance a standardized version of $T_j$ is compared a normal quantile
This yields a test with type I error that is far too large.  An exception is the discussion by Buja et. al. of \citet{LTTT2014}, who
describe their ``max-t'  test that uses permutations at each step to get proper control of type I error.

The conditional max-t test uses the same form of statistic of the max-t test of  described by Buja et. al in \cite{LTTT2014}.
Further, at the first step of forward stepwise (test of the global null), the tests are the same.
They both use parametric or nonparametric sampling of the data to estimate p-values.
it is at subsequent steps that they differ. The max-t test computes residuals from the current model and then applies the global test
to these residuals. In the process, it ignores the selection that has taken place up to the current step. The
conditional max-t test conditions on this selection and uses resamples conditional on this selection.

Because the max-t test ignore the effect of selection, it will tend to be too conservative and lose power in the process.
 Figure \ref{fig:buja} demonstrates this effect in a small example with  three  signal
variables. 
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{bujaplot.pdf}  
\caption{\em Simulated data: $n=20, p=10, \sigma=1$, $\beta=(2,2,2,0,\ldots 0)$. Shown are the p-values for the first 6 steps of forward stepwise regression over 2000 realizations.
In steps 4-6, we only report p-values for truly null case, that is, realizations for which the first 3 predictors are chosen in the first 3 steps.
The power of the max-t and conditional max-t tests (parametric residual  bootstrap and its conditional version) at $\alpha=.10$. are indicated in the bottom right of each of the first 3 panels.}
\label{fig:buja}
\end{figure}
The details are in the Figure caption. We see that the tests are identical within numerical simulation error in the first step, but the max-t test has lower power
in steps 2 and 3.  After that, it becomes more and more conservative.

\section{Optimal tests after forward selection}
\subsection{Hypothesis testing in regression}
\citet{fdrlasso} present a clear definition of the different kinds of hypothesis tests that are possible in stepwise regression.
We summarize their definitions here, as they provide an important framework for our comparisons.
Suppose that we have a ${n \times p}$ data  matrix $X$  and an outcome vector 
$y$, and seek to fit the linear regression model
$$ \hat y =X \beta ^*$$
using a sparse weight vector $\beta^*$. Forward stepwise regression yields a sequence of nested models
$$ \emptyset = \mm_0 \subset \mm_1 \subset ... \subset \mm_p, \; \text{with} \; \mm_k = \{j_1, \, ..., \, j_k\}. $$
We wish to  pick one of the models $\mm_k$, and
set to zero all coordinates $\beta^*_j$ with $j \notin \mm_k$.
The $k$-th ordered hypothesis $H_k$ then tests whether
or not adding the $k$-th variable $j_k$ was helpful or not.

The null hypothesis $H_k$ that the addition of the  $k$-th variable
along the regression path was unimportant, can be formalized
in several ways:
\begin{itemize}
\item {\bf The Incremental Null:} Here we construct $H_k$ to measure whether model $\mm_k$ improves over $\mm_{k-1}$.
The null hypothesis states that the best regression
fit for model $\mm_{k-1}$ is the same as the best regression fit for $\mm_k$ or,
more formally, that
\begin{align}
\label{eq:incremental}
&H_k^{\inc}: \pp_{\mm_{k - 1}} X \beta^* = 
\pp_{\mm_{k}} X \beta^*, \; \text{where} \\
&\pp_{\mm} = X_{\mm} \left(X_{\mm}^\top X_{\mm}\right)^{\dagger} X_{\mm}
\end{align}
is a projection onto the column-span of $X_\mm$. Here, we
write $X_{\mm}$ for the matrix comprised of the columns of $X$ contained
in $\mm$, and $A^\dagger$ denotes the Moore-Penrose pseudoinverse of 
a matrix A. \citet{TLTT2014} develop tests for $H_k^\inc$ in the context of
both forward stepwise regression and least-angle regression.
\item {\bf The Complete Null:}  We consider the stronger null hypothesis
that the model $\mm_{k - 1}$  captures all the available signal. More specifically,
writing $\mm^*$ for the support set of $\beta^*$, we define
\begin{equation}
\label{eq:complete}
H_k^\comp: \mm^* \subseteq \mm_{k - 1}.
\end{equation} 
Tests of $H_k^\comp$ for various pathwise regression models
have been studied by, among others by \citet{LTTT2013} and \cite{LT2014}.
\item {\bf The Full-Model Null:} Here we test
\begin{equation}
\label{eq:abs}
H_k^\FM: j_k \in \mm^*,
\end{equation}
i.e., that the $k$-th variable added to the regression path belongs to the support set of
$\beta^*$. Despite its simple appearance, however, the hypothesis $H_k^\FM$ is
difficult to work with. The problem is that the truth of $H_k^\FM$ depends critically on variables
that may not be contained in $\mm_k$, and so $H_k^\FM$ will have a ``high-dimensional''
character even when $k$ is small. We do not consider the
Full-model Null in this  paper.
\end{itemize}
As we will make clear, the conditional max-t test is designed for testing the Complete Null hypothesis,  while the spacing test of \citet{TLTT2014}
is designed for the Incremental null setting. However the simplicity of the spacing test makes it a natural choice for the Complete Null hypothesis yes as well,
and we examine its performance  in that problem.
\subsection{The conditional max-t test}

To define our conditional max-t test and contrast it with other tests,  we need to make precise the notion of selection.
Suppose at a given step we have selected predictors (and their signs) $S_M$ before this step $(M$ for model) , and predictor (and its sign) $S_C$ at the current step $(C$ for current).
Let the support of the true model be $\cS$.
We test the complete null hypothesis
\begin{eqnarray}
H_0: \cS \subseteq M.
\end{eqnarray}
That is, the null hypothesis is that all signal predictors are contained in the model $M$, and all predictors not in $M$ have coefficients of zero.
Then the conditional max-t test  provides a cutpoint $c$ so that the selective type I error is equal to a target level $\alpha$:
\begin{eqnarray}
\Prob_{H_0} (|T_j| \geq t | S_M)=\alpha
\label{eqn:null}
\end{eqnarray}
More precisely,  $S_M$ is the set of $y$ that lead to selection of the same model $M$ and its signs, in the steps leading up to the current one.

Taking $Z$ to be an $n$-vector of standard normal variates , there are three probability mechanisms that we consider for the probability in (\ref{eqn:null}):
\begin{eqnarray}
y^*=P_My+ \sigma Z;  \;{\rm ``residual} \;{\rm  bootstrap}\;{\rm (RB)''}
\label{eqn:parboot}
\end{eqnarray}
and
\begin{eqnarray}
y^*&=&P_My+ \sigma (I-P_M) Z;  \;{\rm ``conditional}\;{\rm  residual} \;{\rm bootstrap (CRB)}- \sigma \;{\rm known''.} \cr
y^*&=&P_My+ k\cdot (I-P_M) Z;  \;{\rm ``conditional}\;{\rm  residual} \;{\rm bootstrap (CRB)} - \sigma \;{\rm unknown ''.}
\label{eqn:selective}
\end{eqnarray}

The first is a standard parametric residual bootstrap sample from the fitted model $P_M y$;
in the second and third model,s the residuals are constrained  to be orthogonal to $P_My$. This is equivalent to
conditioning $y^*$ on  the observed value of $P_My$, and is the test for the complete null hypothesis that is  derived from the selective framework of \citet{FST2014}.
As a result, under the reduced model it has exact type-I error control.  In the last model above, $k$ is chosen so that $||y^*||= ||y||$.
This corresponds to conditioning on the value of $||y||$.

For the test of the global null hypothesis, \citet{ACP2012}  show that under sparse alternatives where $p^{1-\gamma}$ coefficients are non-zero, and $3/4 \leq \gamma \leq 1$,
the max-t  test is asymptotically most powerful. This result does not address the selective  setup: that is, further steps of the forward stepwise procedure.

Hence we consider the alternative
\begin{eqnarray}
H_1: p^{1-\gamma}\;{\rm predictors}\; {\rm not}\;{\rm in}\; M\;{\rm  have}\; {\rm  nonzero}\;{\rm coefficients}
\end{eqnarray}
with $3/4 \leq \gamma \leq 1$.
Building on the result of \citet{ACP2012}, heuristically we would expect that  the conditional max-t test is the asymptotically most powerful selective test for this problem under appropriate conditions.
\subsection{Summary of the various tests studied}

Since a number of different tests are examined in this paper, we summarized them here.
\begin{enumerate}
\item {\em Max-t test} with Gaussian errors (\ref{eq:mod}) . This test does not condition on the selection done in previous steps.
\item {\em Max-t-RB test}. The conditional max-t test using bootstrap  sampling from the  model (\ref{eqn:parboot}).  This is done either parametrically (sampling from the Gaussian)
or non-parametrically (sampling from the model residuals).  The parametric sampling is discussed in Section \ref{sec:exact}, while 
the nonparametric version is illustrated in Section \ref{sec:bootstrap}.
\item {\em Max-t-RB approximate test}. 
 This is an analytic approximation to the  parametric bootstrap Max-t p-value, given below in Section (\ref{sec:analytic}).
\item {\em Max-t-CRB test}. The conditional max-t test using bootstrap  sampling from the reduced model as in (\ref{eqn:selective}).  This is done either parametrically (sampling from the Gaussian)
or non-parametrically (sampling from the model residuals).  T
\end{enumerate}

\section{Computation of the conditional max-t test p-values}

\subsection{Exact computation}
\label{sec:exact}
Given a selected model $M$, we need to estimate the probability in (\ref{eqn:null}).
For the first step of forward stepwise (for which  $M$  is empty), this is straightforward:  we can fix $X$ and generate permutations (or bootstrap samples)
of $y$, compute the corresponding test statistics $T_1^*, T_2^*, \ldots T_p^*$  and use this to estimate the tail probability directly.

For subsequent steps of
forward stepwise regression, the computation is much more challenging. Given a model $M$ with fit $\hat\mu_M$,  we  generate a bootstrap sample
\begin{eqnarray}
y_i^*=\hat\mu_i+ \sigma A \cdot Z_i, \; i=1,2,\ldots N
\label{eq:boot}
\end{eqnarray}
Here $A$ equals the identity for the parametric bootstrap test, and $(I-P_M)$ for the selective test.
(Alternatively we can do residual bootstrap sampling: see Section \ref{sec:bootstrap}) .
 Then we can directly estimate the tail probability in  (\ref{eqn:null}). However we need to retain
only realizations $y^*$ that fall in $S_M$, that is, yield the same selected model $M$.

The task is made much simpler by the fact that the selection event can we written as $S_M=\{y: \Gamma y \geq u\}$.
Hence we don't need to run the entire forward stepwise procedure on $y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check
if $\Gamma y^* \geq u$. Further, since $\Gamma$ has on the  order of $2pk$ rows after $k$ steps,  we can save 
substantially in computation by checking the row-wise inequalities starting from the last row, wherethe inequality  it is most likely to fail.

An additional fact that makes the computation feasible with strong signals:  much  of the time the procedure applied to bootstrap samples
make the same selections as in the original sample. It is only when the selected variables are  noise that the great majority of bootstrap samples
are rejected. Hence we run the procedure for a few steps after  the estimated p-values start to rise, and then stop.

 {\em Importance sampling} is also very helpful here.
Suppose that we want to estimate the conditional cdf  of a statistic $T(y)$ under $N(\mu, \Sigma)$, denoted by
$F_\mu(t \mid S)$.
Then if we instead sample $y_1^*, \ldots y_B^*$ from $N(\mu_0,\Sigma)$,  giving estimates $T_1^*, T_2^*, \ldots T_b^*$, 
an unbiased estimate of $F_\mu(t)$ is given by
\begin{eqnarray}
\frac{1}{B} \sum_b w_b \cdot I(T_b^* \leq t)
\end{eqnarray}
where $w_b=\phi_{\mu,\Sigma}(y_b^*)/\phi_{\mu_0,\Sigma}(y_b^*)$.
The idea is that we can choose $\mu_0$ so that the probability of $y^*$  lying  in the selection set $S$ is larger than under
$\mu$ (ie. fewer rejections).

Here are more details.
We want to choose $\mu_0$ to decrease the proportion of rejections, but don't want $\mu_0$ to be too far from $\mu$, or else the weights
will become too extreme.
 Denote the least squares estimate for the selected model after $k$ steps by $\hat\beta^M$.
We construct $\mu_0$ as follows.
The expected maximum of $(p-k)$  $N(0,1)$ variates is $\sqrt{2\log(p-k)}$.
Thus for  each component $j$, if $|\hat\beta_j^M/{\rm \hat se}_j|< \sqrt{2\log(p-k)}$ we increase $\hat\beta_j^M$ away from zero 
so that it equals $\sqrt{2\log(p-k)}$. For signal predictors, the coefficient will usually be unchanged: but for noise predictors, this will inflate their
coefficient so that there will be fewer rejections.

Table \ref{tab:tab1} shows an example. For the most difficult  case ($p=30$,  8 signals, 10 steps) the procedure required  1.8M samples to obtain 300 good ones.

\begin{table}[ht]
\centering
\begin{tabular}{r|rrr}
  \hline
  p&\multicolumn{3}{c}{Number of signals}\\
 & 2 & 5 & 8 \\ 
  \hline
  10 & 0.08 & 1.62 & 1.23 \\ 
  20 & 0.10 & 1.11 & 14.90\\ 
  30& 0.09 & 0.35 & 24.81 \\ 
   \hline
\end{tabular}
\label{tab:tab1}
\caption{\em Time in seconds to estimate p-values, for nsteps= \# signals+2; min 300 bootstraps; n=50}
\end{table}

\subsection{An analytic approximation}
\label{sec:analytic}
In this section we provide an analytic approximation to the parametric bootstrap (\ref{eqn:parboot}).
First consider the global null case, and let $T_j=v_j^Ty$ be the least squares estimates for predictor $j$  and $T_{max}={\rm max}_j(|T_j|)$ with observed value $M$.
Let $G(t)$ be the CDF of $T_{max}$. The desired p-value is $1-G(M)$.
Recall  Sidak's  lemma (see e.g. \citet{tong1990}) which states that if $y \sim N(0,\Sigma)$ then
\begin{eqnarray}
\Pr(\cap\{ |y_i| \leq t_i\} \geq \prod \Pr(|y_i| \leq t_i)
\end{eqnarray}
with equality if the variables are independent. Using this fact, we can approximate the p-value  $1-G(M)$ by 
$1-\prod \Pr(|T_j| \leq M)$ and this will be conservative by Sidak's lemma.

Now after  $k$ steps of FS with $p$ predictors, denote the selection by $S$. Let $T_j=v_j^Ty$,  $T_{max}={\rm max}_j(|T_j|)$
with observed value $M$. Now at the first step, the value of $T_{\rm max}$ can be any value in $(-\infty, \infty)$.  But after $k$ steps, the selection  events allow us to narrow this range.
The simple max-t test  described by Buja et. al in \cite{LTTT2014} ignores this fact and uses the range $(-\infty, \infty)$ for all steps, which will make quite conservative (as seen in Figure \ref{eqn:bujaplot}).
The narrowing of the range is most easily seen when the predictors are uncorrelated. 
Then $|T_{max}|$ can be no larger than its value at the previous step.  Hence $G(m)$ is the  CDF of largest of the last $p - k + 1$ 
absolute normal order statistics, conditioned on being smaller than the $k$th  order statistic.

Consider then the general non-independent  case where we have made a selection $S$ after $k$ steps.
Let predictor $0$ be the one selected at the current step.
Let $G(t)$ be the CDF of $T_{max} \mid S$. 
Let $V_0^\pm$ be the limits derived from polyhedral lemma applied to $S$ and $v_0$.
This lemma is reviewed in the appendix, and allows us to write the selection set $y \in S$  as
$V_0^- \leq v_0^Ty\leq V_0^+$.

Let $V={\rm max}(|V_0^-|, |V_0^+|)$, $M'=min(V,M)$.
Define
\begin{eqnarray}
F(M')=\Bigl[\frac{\Phi(M'/\sigma)-\Phi(-M'/\sigma)}{\Phi(V/\sigma)-\Phi(-V/\sigma)}\Bigr]^{p-k}.
\label{eq:approx}
\end{eqnarray}
The quantity $1-F(M')$ is our proposed approximate p-value.
We conjecture that $G(M') \geq F(M)$  so that $1-F(M')  \geq 1-G(M) \sim U(0,1)$ under the null. Hence the test based on $1-F(M')$ is conservative under the null.
At the first step (global null)  we have $V=\infty$ and hence this result is already established above. For subsequent steps with independent predictors, it is easy show that 
$V$ equals the $k$th  largest value among the $|T_j|$ and again the result follows.

Now we sketch a proof of the general result.
First we give  a refinement of  Sidak's lemma, which we need later.

\medskip

{\bf Lemma 1.} 
Let $y \sim N(0,\Sigma)$ and define
$$S(t) = \frac{\Pr({\rm max} |y_i| \leq t )}{\prod_i \Pr( |y_i| \leq  t )},$$
Then $S(t)$ is non-increasing in $t$.  [We note that Sidak's lemma implies that $S(t) \geq 1= S(\infty)$.]
A proof, due to Roman Latala is given in the Appendix. We were unable to find the result in the literature.

Now consider $G(M \mid S)=\Pr({\rm max}_{j \notin S} (|T_j|\leq M \mid S)$.

PROOF IS NOT YET CORRECT!

We can write
\begin{eqnarray}
G(M \mid S, V_0^- ,V_0^+)&=&\Pr({\rm max}_{j \notin S} (|T_j|\leq M\mid V_0^- \leq v_0^Ty\leq V_0^+,V_0^- ,V_0^+)\\
&\geq &\Pr({\rm max}_{j \notin S} (|T_j|\leq M\mid -V \leq v_0^Ty\leq V, V_0^- ,V_0^+) )\\
&=&\Pr({\rm max}_{j \notin S} (|T_j|\leq M\mid  -V \leq v_j^Ty\leq V, j\notin S, V_0^- ,V_0^+)) \\
&\geq &\Pr({\rm max}_{j \notin S} (|T_j|\leq M'\mid -V \leq v_j^Ty\leq V, j\notin S, V_0^- ,V_0^+)\\
&=&  \frac{ {\Pr({\rm max}_{j \notin S} |T_j|\leq M'},V_0^- ,V_0^+) } { \Pr({\rm max}_{j \notin S} |T_j|\leq V, V_0^- ,V_0^+)}\\
&\geq & \frac{\Bigl[\prod_{j\notin S} \Pr(|T_j|\leq M', V_0^- ,V_0^+)\Bigr] } { \prod_{j\notin S} \Pr(|T_j|\leq V, V_0^- ,V_0^+]}\\
&=&\Bigl[\frac{\Phi(M'/\sigma)-\Phi(-M'/\sigma)}{\Phi(V/\sigma)-\Phi(-V/\sigma)}\Bigr]^{p-k} \\
\end{eqnarray}
Now $V_0^- ,V_0^+$ are independent of $v_j^Ty$ for $j\notin S$, so (13) holds if we integrate out wrt $V_0^- ,V_0^+$.

Explanations:

(14):  $|V| \geq |V_0^-|, |V_0^+|$.

(15)  predictor 0 was selected among $j \notin S$

(16)  $M' \leq M$.

(18)  follows from Lemma 1

(19) The variance of $v_j^Ty$ is $\leq \sigma$ and under the null each $T_j$ has mean zero.

\subsection{Example}

Figure \ref{fig:fig2} compares the spacing, exact and approximate conditional max-t tests in  two examples.
We see that conditional max-t test has substantially more power. The exact conditional max-tmax-t test test becomes
a little conservative in the top  right panel, as we move farther into the null. The reason is as follows.
Let $\hat \mu_0$ by the least squares fit for the model containing just the signal variables, and suppose that we have fit a larger model
$\hat\mu$ containing the signal variables plus some noise ones.  Then the correct procedure generates
$y^*=\hat\mu_0+\sigma Z$, but instead we generate
$$y'=\hat\mu+\sigma Z=y^*+(\hat\mu-\hat\mu_0)$$
Consider a contrast vector $v$ that is orthogonal to the column spaces yielding both $\hat\mu_0$ and $\hat\mu$.
Then 
$$v^T y'=v^T y^*+v^T(\hat\mu-\hat\mu_0)$$
The second term above in normal  mean zero and some variance $\tau^2$.
Thus the variance of $y'$ is larger than that of $y^*$ and the test becomes conservative.

\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{fig2.pdf}  
\caption{\em $n=20, p=8$ ,  $\beta=(3,0,0\ldots)$ in top panels and $(3,3,0,0,\ldots)$ in bottom panels.
Estimated power at $\alpha=.10$ for spacing, approximate and exact conditional max-t tests: top left- .60, .80. 80;  bottom left- .59, .98,.98,; bottom middle-
.39, .56, .62. }
\label{fig:fig2}
\end{figure}

\subsection{Relation to the spacing test} 
\label{sec:spacing}
Suppose that we have carried out a forward selection  yielding a set of predictors and their signs, denoted by $S$.
Let $\theta={\rm E}(y)$ and suppose we wish to make inferences about some linear contrast $v^T \theta$ based on
our sample estimate $v^T y$. The common choice for $v$ is the vector that produces the partial regression coefficient for one of the predictors
in the chosen model. Note that $v$ is data-dependent.
\citet{TLTT2014} and \citet{LT2014} propose a closed form method that yields exact tests for $H_0: v^T\theta=0$ and exact confidence intervals
for$v^T\theta$, condition on the selection $S$.
From this description, we see that the spacing test is a test of the incremental null hypothesis.

\cite{FST2014} provide a general framework for inference after selection, and derive uniformly most tests  when the distribution of the data are in  the exponential family.
There are two scenarios in their paper: the {\em saturated model} for which only model ({\ref{eq:mod}) is assumed, and the {\em reduced model} in which we assume
that some submodel of ({\ref{eq:mod}) holds. \citet{FST2014} 
show that when $\sigma$ is known, the spacing test coincides with the uniformly most powerful test in the saturated model. Furthermore. if the predictors are orthogonal
the spacing test is optimal under the reduced model as well.

There is another conceptual difference between the two tests.
If we include the current selection, we can view the overall selection event as $S=\{S_P,S_c\}$. The spacing test conditions on all of $S$.
The conditional max-t test conditions only on $S_M$. Less conditioning  generally yields more power.

The following simple example contrast the two tests explicitly,
Suppose we have $y=(y_1,y_2) \sim N((\theta_1,\theta_2), I)$.
Assume that we select the variable $y_j={\rm max}(y_1,y_2)$.
If we observe $y_1>y_2$, the spacing test  uses  the null hypothesis $\theta_1=0$, and its p-value is easily shown to be
\begin{equation}
{\rm pv}_{\rm spacing}=\frac{1-\Phi(y_1)}{1-\Phi(y_2)}.
\end{equation}

On the other hand, the conditional max-t test uses  the null hypothesis $\theta=(0,0)$ and has p-value
\begin{equation}
{\rm pv}_{\rm max}=1-\Phi(y_1)^2
\end{equation}
Direct inspection shows that ${\rm pv}_{\rm max} < {\rm pv}_{\rm spacing}$  as long as $y_1>y_2 > 0$.

As a second example, \citet{TLTT2014} derive a sequential test  (``the spacing test'') for the Gaussian regression model, with an exact analytic formula for the p-values. This is one of the first
regression tests that properly account for the selection effects and has exact type I error control at each step.

The left panel of Figure \ref{fig:fig1}  shows a simple example of the conditional max-t test compared to the spacing test.
 There are just two orthogonal predictors, whose statistics $T_1, T_2 \sim N(\Delta,1)$.
 We select the largest of the two statistics in absolute value. The power of the conditional max-t test , for target type I error $\alpha=0.10$, is shown by the red curve, as a function of $\mu$.
  Note that at $\Delta=0$ the power is $0.10$, confirming that the test is properly calibrated,
In Figure \ref{fig:fig1} we can see that the spacing test has substantially less power in this example.
The case of two large equal effects turns out to be especially problematic for the spacing test. In the right panel we show the alternative $(\Delta, \Delta/2)$:
the gap between the power curves has narrowed, but the conditional max-t test is still superior.
We study the relationship of the conditional max-t test with the spacing test in Section \ref{sec:spacing}.

\begin{figure}[hbtp]
\centering
\includegraphics[width=.6\textwidth]{diab.pdf}  
\caption{\em Diabetes data: p-values from spacing test, and exact and approximate conditional max-t test}
\label{fig:diab}
\end{figure}

\subsection{Robustness to model misspecification}
Figure \ref{fig:bujaplot-unksig} shows the sequential p-values from a setting in which the reduced is misspecified.
Details are in the Figure caption.  We wee that the type I error of the conditional CRB test  started to drug above the nominal level in steps 
4,5, and 6.
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{bujaplot-fixnormy-unksigma.pdf}  
\caption{\em $n=20, p=8, \beta=c(4,3,2,rep(0,5),rep(1,8))$, pairwise correlation $0.5$; only $x_1, \ldots x_8$ are
observed; $\sigma$ unknown}
\label{fig:bujaplot-unknsig}
\end{figure}

\section{Diabetes data}
Here we apply the various selection procedure to some well-known known on diabetes.
These data consists of  observations on $442$ patients, with the response of interest being a quantitative measure of disease progression one year after baseline.
There are ten baseline variables---age, sex, body-mass index, average blood pressure,
and six blood serum measurements---plus quadratic terms, giving a total of 64 features.
\begin{figure}[hbtp]
\centering
\includegraphics[width=.6\textwidth]{diab.pdf}  
\caption{\em Diabetes data: p-values from spacing test, and exact and approximate conditional max-t test}
\label{fig:diab}
\end{figure}
The results are shown in Figure \ref{fig:diab}. We see that there are only 2 or 3 small spacing p-values, while both forms of the conditional max exhibit 5 small p-values.

\section{A nonparametric conditional max-t test }
\label{sec:bootstrap}
The non-parametric version of the conditional max-t test simply replaces the Gaussian generative model (\ref{eq:boot}) by bootstrap residual sampling:
\begin{eqnarray}
y_i^*=\hat\mu_i+ r_i^*,\; i=1,2,\ldots N
\label{eq:bootnp}
\end{eqnarray}
where $r_i*$ is sampled with replacement from the residuals $r_i=y_i-\hat\mu_i$.
We then apply the same exact computation scheme as outlined above.
By avoiding normality assumption for the errors,  the hope is that this procedure will be more robust that the parametric version.
Figure \ref{fig:slash} examines this in a small example with no signal.
In the top row, with Gaussian errors, we see that both methods provide type I error control.
In the bottom two, we have used long-tailed ``slash'' errors (a Gaussian divided by a uniform), and we use that the parametric  test becomes
badly anti-conservative.  However the nonparametric version holds up quite well.
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{slash.pdf}  
\caption{\em Parametric and nonparametric bootstrapping;. $n= 15, p=5$. Null model.
Gaussian or slash errors}
\label{fig:slash}
\end{figure}

\subsection{Analytic approximation to the nonparametric bootstrap test}
Given a selection $S$, let $G(M \mid S)=\Pr({\rm max}_{j \notin S} (|T^*_j|\leq M \mid S)$.
Let $V$ be the bounds  in (\ref{eq:approx}) obtained from polyhedral lemma.
Then the same style of  argument that  was used in the Gaussian case suggests the approximation
\begin{eqnarray}
G(M \mid S, V_0^- ,V_0^+) &\approx& \Pr({\rm max}_{j \notin S} (|T^*_j|\leq M\mid V_0(y)^- \leq v_0^Ty^*\leq V_0(y)^+,V_0(y)^- ,V_0(y)^+)\cr
&=& \frac{ {\hat F}(M)-{\hat F}(-M)} {  {\hat F}(V)-{\hat F}(V)}
\label{eq:bootapprox}
\end{eqnarray}
Here $\hat F$ is the bootstrap distribution of the maximum attained  at step $k$. No rejection sampling is needed for account for the selection.
Figure \ref{fig2boot} suggests that this approximation is potentially useful.

\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{fig2boot.pdf}  
\caption{\em Approximation to non-parametric residual bootstrap p-values using (\ref{eq:bootapprox}). Here  $n=20, p=8$ and  $\beta=(3,0,0\ldots)$ in top panels and $(3,3,0,0,\ldots)$ in bottom panels.}
\label{fig:fig2boot}
\end{figure}

{\bf Proof of correctness of bootstrap procedure}

\section{Other topics and issues}
Here we briefly discuss some further topics.
\begin{itemize}
\item {\it Estimation of the noise variance}
\item{\it Backward stepwise selection}.
The method described here can be also be adapted to backward stepwise regression.
Here are some details.
Suppose that we have dropped some  $k$ predictors: call the drop event $D$.
Consider another  predictor  $j$ as a candidate to drop. Let $r=P_{D-j}^\perp y$, $\tilde X=P_{D-j}^\perp X$,
$W_j=\langle r, \tilde x_j\rangle/||\tilde x_j||$. Then we drop the predictor
\begin{eqnarray}
\hat j={\rm argmin }(|W_j|)
\end{eqnarray}
Denote the attained minimum by $M=W_{\hat j}$. Define
$$G(t)=P(M\geq t | D) $$
then our p-value is $1-G(W_{\rm min})$.
We can use the same form of construction as before to get an analytic approximation for $G$:
\begin{eqnarray}
F(M')=\Bigl[\frac{ (1-\Phi(M'/\sigma)-\Phi(-M'/\sigma))}{(1-\Phi(V/\sigma)-\Phi(-V/\sigma))}\Bigr]^{p-k}\
\label{eq:approx}
\end{eqnarray}
The quantity $1-F(M')$ is then our proposed approximate p-value.
\end{itemize}

\section{Results of various simulations}

\subsection*{Acknowledgements}

\appendix
\section*{Review of Conditional Gaussian inference after polyhedral selection}

This section describes  a powerful result on Gaussian contrasts after polyhedral selection, which
appears in \citet{TLTT2014} and \citet{exactlasso}. 
\begin{equation*}
y \sim N(\theta,\Sigma), 
\end{equation*}
where $\theta \in \R^n$ is unknown but $\Sigma \in \R^{n\times n}$ 
is known. This generalizes our original setup in \eqref{eq:mod} (in
that we allow for a general error covariance structure).  Consider a 
polyhedron 
\begin{equation*}
\cP = \{y : \Gamma y \geq u\},
\end{equation*}
where $\Gamma \in \R^{m\times n}$, $u \in \R^m$ are fixed, and the
above inequality is meant to be interpreted componentwise.  Recall
that we call $y\in\cP$ a polyhedral selection event. Our goal is to
make inferences about $v^T \theta$ conditional on $y \in \cP$.
Indeed $v=v(\cP)$ is allowed to be defined in terms of $\cP$.  The 
next result provides a key alternate representation for
$\cP$.

It is shown in \citet{TLTT2014} and \citet{exactlasso}.  that
for any $\Sigma,v$ such that $v^T \Sigma v \not= 0$, there exists explicitly computable
functions $\V^\mathrm{lo}(y), \V^\mathrm{lo}(y), \V^0(y)$ such that
\begin{equation}
\label{eq:truncate}
\Gamma y \geq u \iff
\V^\mathrm{lo}(y) \leq v^T y \leq \V^\mathrm{up}(y), 
\, \V^0(y) \leq 0,
\end{equation}
Moreover,  it turns out that the triplet $(\V^\mathrm{lo},\V^\mathrm{up},\V^0)(y)$ is
independent of $v^T y$.    The explicit  expressions for these quantities are given  
 in the Appendix.

Hence the distribution of 
any linear function $v^Ty$, conditional on the selection 
$\Gamma y \geq u$,  
can be written as a truncated Gaussian distribution, and this can be used for hypothesis testing and confidence interval
construction. 
\citet{TLTT2014} use this result to derive a test for the entry of predictors in least angle and forward stepwise regression.

\section*{ Proof of Lemma 1}  {\em Due to Rafal Latala, personal communication}

The standard proof of the Khatri-Sidak inequality  (e.g. \citet{J70}) shows that for any
symmetric convex set $K$ and symmetric Gaussian vector $X$ ,the function

$$t \rightarrow  \frac {\Pr(X\in K,|X_1|< t)}{\Pr(|X_1|<t)}$$
is nonincreasing on $(0,\infty)$.
This shows that the function
$$G(t_1,\ldots,t_n)=\frac{\Pr(|X_1|< t_1,\ldots,|X_n|<t)}{\prod_i \Pr(|X_i|<t_i)}$$
is coordinate-wise nonincreasing on$ (0,\infty)^n$, in particular
$S(t)=G(t,\ldots,t)$
is nonincreasing on $(0,\infty)$.

\bibliographystyle{agsm}
\bibliography{ryantibs}

\end{document}