### 2015-05-22 notes/nonparam.tex
notes on nonparametric bootstrap

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\newcommand{\real}{\mathbb{R}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\title{Nonparametric selective bootstrap}

\maketitle

\section{Non-bootstrap world}

Suppose we have a model selection procedure $\hat{M}$ acting on sufficient
statistics $Z$ choosing among a (finite) set of models. So, our selected model
is $M=\hat{M}(Z)$. Our selected model suggests a hypothesis of interest that we want to test.

We will apply $\hat{M}$ on data of the form
$$
\hat{Z}_n = \frac{1}{n} \sum_{i=1}^n y_i X_i = \frac{1}{n}X^Ty \in \real^p
$$
where $(y_i, X_i) \overset{\text{IID}}{\sim} F$.

Note that $Z$ is a statistical functional that we might write as
$$
Z(F) = E_{F}(yX)
$$
and what we really observe $\hat{Z}_n = Z(\hat{F}_n)$.

We want to carry out a selective test of
$$
H_0: \eta_{M}^T \gamma_M(F) = 0
$$
where $\eta_M \in \real^{p \times k}$ and $\gamma_M$ is some statistical
function possibly different from $Z$. In principle it can depend
on the model we have selected, but we will drop the $M$ from now on.
{\bf We assume that $Z$ is measurable with respect to $\gamma$, i.e.
$\gamma$ includes $Z$ as part of its coordinates or something. This means
that it makes sense to talk about $\hat{M}$ evaluated
on a sample of $\gamma$'s.}

Above, we have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
selected model framework, we just have to design
a test for a fixed $M$.

For large enough $n$, under our model $M$ suppose that
$$
\gamma(\hat{F}_n) \approx N(\gamma(F), n^{-1} \Sigma(F))
$$
where, for some distribution $G$ on $\real \times \real^p$
$$
\Sigma(G) = E_{G}\left((\gamma(\tilde{y},\tilde{X}) - \gamma(G))(\gamma(\tilde{y},\tilde{X}) - \gamma(G))^T \right)
$$
where $(\tilde{y}, \tilde{X})$ is a sample of size 1 from $G$.

{\bf Note: as the functional
we choose may depend on $M$, the normal approximation generally depends 
on the selected model -- I am not going to emphasize that here.}

If this is our reference distribution, we should base our distribution on the 
conditional law of $\Sigma(F)^{-1}W | (I - \eta\eta^{\dagger})\Sigma(F)^{-1}W$ where
$W \sim N(\gamma(F), n^{-1} \Sigma(F))$. This is because
we know that, under $H_0:\eta^T\gamma(F)=0$
this distribution does not depend on $(I-\eta\eta^{\dagger})\gamma(F)$ which
is everything except what we wanted to test.

Set $P_{\eta}^{\perp}=I - \eta\eta^{\dagger}$.
This 
distribution has covariance
$$
n^{-1} \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right).
$$
Denote this covariance by $\Theta(\eta, \Sigma(F))$.

If $W \sim N(\gamma(F), n^{-1}\Sigma(F))$,
this distribution can be realized by 
$$
\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
\Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W.
$$
This is an affine function of $W$ determined by $\eta, \Sigma(F)$ 
call this affine function $A(\eta, \Sigma(F), W)$.

In summary, if $W \sim N(\gamma(F), n^{-1}\Sigma(F))$ 
then under $H_0: \eta^T\gamma(F)=0$, asymptotically we have
$$
A(\eta, \Sigma(F), W) \sim N(0, n^{-1} \Theta(\eta, \Sigma(F))).
$$

{\bf Note that, if we can
sample from this distribution, we know how to
do a parametric selective test of $H_0:\eta_M^T\gamma(F)=0$.
Also, I reiterate that the normal approximation could depend on $M$.
}

\section{Bootstrap test}

We apply non-parametric bootstrap to the pairs 
yielding samples
$$
y^{*,b} \in \real^n,  X^{*,b} \in \real^{n \times p}, \qquad 1 \leq  b \leq B
$$
and  bootstrapped sufficient statistics
$$
\gamma^{*,b} = \frac{1}{n} \gamma(y_i^{*,b}, X_i^{*,b}),  \qquad 1 \leq  b \leq B
$$

These are of course just evaluating $\gamma$ on 
IID draws of size $n$ from $\hat{F}_n$. So,
we can write
$\gamma^{*,b} = \gamma(\hat{F}_n^{*,b})$
where $\hat{F}_n^{*,b}$ is the empirical
distribution function of sampling with replacement from the original data.

For large enough $n$, the law of $\gamma(\hat{F}_n^{*,b})$ is hopefully going to be close
to the Gaussian 
$$
\gamma(\hat{F}_n^{*,b}) | \hat{F}_n \sim N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))
$$
where
$$
\Sigma(\hat{F}_n) = \frac{1}{n} \sum_{i=1}^n \left(\gamma(y_i,X_i) - \bar{\gamma} \right)\left(\gamma(y_i,X_i) - \bar{\gamma} \right)^T
$$
and
$$
\bar{\gamma} = \frac{1}{n} \sum_{i=1}^n \gamma(y_i, X_i).
$$

(Again, the normal approximation may depend on the selected model.)

Therefore, our unselective bootstrap should be giving samples from
roughly $ N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))$. That is,
taking $B$ very large:
$$
\frac{1}{B}\sum_{b=1}^B \eta^T \gamma(\hat{F}^{*,b}_n) \approx \eta^T\gamma(\hat{F}_n).
$$
This suggests we correct the unselective bootstrap by drawing
$$
\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n) | \hat{F}_n \approx N((I - \eta\eta^{\dagger})\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n))
$$
and we see that
$$
A(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger} \gamma(\hat{F}_n)) | \hat{F}_n
 \approx N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
$$
As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. In general,
plugging in consistent estimates of variance is OK in selective inference (not sure
if Xiaoying put it in the first paper, but definitely in the
randomized setting we have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
I think it will be fine. Maybe a degrees of freedom correction will help.

So,
we draw nonparametric bootstrap samples
$$
\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n)
$$
keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.
For those that survive, we compute
$$
A(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n))
$$
and take a test statistic of our choosing.

\section{No nuisance parameters}

When there are no nuisance parameters (i.e. $\eta=I$ or we are make other assumptions about $Z(F)$), we do not need the
function $A$. This is the example I tried this morning. In this
case, $\gamma=Z$ and we just draw nonparametric bootstrap samples
$$
Z(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}Z(\hat{F}_n)
$$
keeping the same ones and compute any test statistic of our 
choosing on these bootstrap samples.

\end{document}