### 2015-11-09 paper.tex
paper.tex

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{thmtools, thm-restate}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{geometry}
\usetikzlibrary{shapes,arrows}
\tikzstyle{block} = [rectangle, draw, fill=white!20,
    text width=7em, text centered, rounded corners, minimum height=4em]
\tikzstyle{title} = [text width=7em, text centered, font=\bfseries]
\tikzstyle{line} = [draw, -latex']

\usepackage{mycommands}

\begin{document}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{model}[theorem]{Model}

\theoremstyle{definition}
\newtheorem{example}{Example}

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
%\newcommand{\cP}{\mathcal{P}}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\supp}{\text{supp}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\tM}{\widetilde{M}}
\newcommand{\tE}{\widetilde{E}}
\newcommand{\tV}{\widetilde{V}}
\newcommand{\tR}{\widetilde{R}}
\newcommand{\tL}{\widetilde{L}}
\newcommand{\hk}{\hat{k}}
\newcommand{\hr}{\hat{r}}       
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
\newcommand{\Err}{\mathcal{E}}
\newcommand{\RSS}{\text{RSS}}

\newcommand*\mystrut{\vrule width0pt height0pt depth1.5ex\relax}
\newcommand{\underlabel}{\underbracket[1pt][.5pt]{\mystrut \quad\;\; \sub \quad\;\; }}
\newcommand{\JTcomment}[1]{{\color{blue}{(JT: \bf \sc #1) }}}
\newcommand{\WFcomment}[1]{{\color{red}{(WF: \bf \sc #1) }}}
\newcommand{\RTcomment}[1]{{\color{green}{(RT: \bf \sc #1) }}}

\title{Selective Sequential Model Selection}
\author{William Fithian, Jonathan Taylor, Rob Tibshirani, and Ryan Tibshirani}
\maketitle

\begin{abstract}
  Many model selection algorithms produce a path of fits specifying a sequence of increasingly complex models. Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data. Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data. In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-$\lambda$ test for the lasso. These tests improve on the power of the spacings test of \citet{taylor2014exact}, sometimes dramatically. In addition, our framework extends beyond linear regression to a much more general class of parametric and nonparametric model selection problems.

To select a model, we can feed our single-step $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate (FWER) or false discovery rate (FDR) as desired. The FDR-controlling rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}. We derive intuitive and general conditions for independence and show that our proposed constructions yield independent $p$-values. We also discuss connections between the selected-model test and parametric bootstrap sampling.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
Many model selection procedures produce a sequence of increasingly complex models, leaving the data analyst to choose among them. 
%While methods like cross-validation aim at minimizing out-of-sample prediction error, such methods tend to select a large fraction of noise variables. 
Given such a path, we consider the problem of choosing the simplest model in the path that is not falsified by the available data. Examples of path algorithms include forward stepwise linear regression, least angle regression (LAR) and  the ever-active path in $\ell_1$-regularized models. \citet{taylor2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals. In addition, \citet{choi2014selecting} discuss a selective approach to estimating the rank of the covariance matrix in principle components analysis.

To select a model, we compute a set of sequential $p$-values at each step of the model path, and then feed them into a stopping rule that is guaranteed to control the false discovery rate (FDR) \citep{benjamini1995controlling}, familywise error rate (FWER), or a similar quantity. Recently \citet{gsell2013sequential} and \citet{li2015accumulation} proposed sequential stopping rules of this kind. Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform and independent. In this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and show how to construct them in many settings.

While our approach and analysis are quite general, we begin by introducing a specific example: the selective max-$t$ test for forward stepwise regression, a selective sequential version of the max-$t$ test of \citet{buja2014}, generalized using the theory in \citet{fithian2014optimal}.

\subsection{The max-$t$ Test in Forward Stepwise Regression}

Forward stepwise regression is a greedy algorithm for building a sequence of nested linear regression models. At each iteration, it augments the current model by including the variable that will minimize the residual sum of squares of the next fitted model. Equivalently, it selects the variable with the largest $t$-statistic, controlling for the variables in the current model.

For design matrix $X\in \R^{n\times p}$ and response vector $Y\in \R^n$, let $E \sub \{1, \ldots, d\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model. For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables and an intercept term. Using the standard result that
\begin{equation}
t_{j,E}^2 = (n-|E|-2) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
\end{equation}
we see that the next variable selected is 
\begin{equation}
j^* = \argmax_{j\notin E} |t_{j,E}| = \argmin_{j\notin E} \RSS(E \cup \{j\}),
\end{equation}
with corresponding $t$-statistic $t_E^* = t_{j^*,E}$. The selective max-$t$ test rejects for large values of $t_E^*$, compared to an appropriate conditional distribution that accounts for the adaptive selection of the model path.

Table~\ref{tab:diab} illustrates the max-$t$ test and two others applied to the diabetes data from \cite{LARS}, consisting of observations of $n=442$ patients. The response of interest is a quantitative measure of disease progression one year after baseline, and there are ten measured predictors --- age, sex, body-mass index, average blood pressure, and six blood serum measurements --- plus quadratic terms, giving a total of $d=64$ features. We apply forward stepwise regression to generate the model path (beginning with an intercept term), and then use each of the three methods to generate $p$-values at each step along the path. Finally, we use the ForwardStop rule \citep{gsell2013sequential} at FDR level $\alpha=0.1$ to select a model based on the sequence of $p$-values. The red entry in each column indicates the last variable selected by ForwardStop.

\begin{table}[ht]
\centering
\input{tables/diabetes_edit.tex}
\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values. The {\bf bold annotation} indicates the model selected by ForwardStop with $\alpha=0.1$. For the max-$t$ $p$-values, this stopping rule gives exact $\text{FDR}_{\text{model}}$ control at the 10\% level.}
\label{tab:diab}
\end{table}

The nominal $p$-values in Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$ distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data. Because the model sequence is in fact selected adaptively to maximize the value of $t_E^*$, this method is highly anti-conservative.

The max-$t$ test $p$-values are computed using the same test statistic $|t_E^*|$, but compared with a more appropriate null distribution. We can simulate the distribution of $|t_E^*|$ under the null model, i.e. the linear regression model specified by active set $E$:
\begin{equation}
  M(E):\; Y \sim N(X_E\beta, \sigma^2 I), \quad \beta\in \R^{|E|}, \sigma^2>0,
\end{equation}
where $X_E$ is the matrix with columns $(X_j)_{j\in E}$. Because $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}. In step one, because $E=\emptyset$ is fixed and $t_E^*$ is independent of $U_E$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}. In later steps, $t_E^*$ and $U_E$ are not conditionally independent given $E$, but we can approximate the null conditional distribution through Monte Carlo sampling.

The saturated-model $p$-values also account for selection, but they rely on somewhat different assumptions, condition on more information, and test a slightly different null hypothesis. We discuss these distinctions in detail in Section~\ref{sec:whichnull}.

For most of the early steps in Table~\ref{tab:diab}, the (invalid) nominal test gives the smallest $p$-values, followed by the max-$t$ test, then the saturated-model test. Both the max-$t$ and saturated-model $p$-values are exactly uniform under the null, but the max-$t$ $p$-values appear to be more powerful under the alternative. As we will discuss in Section~\ref{sec:selective-reg}, selected-model tests such as the max-$t$ can be much more powerful than saturated-model tests in early steps of the path when multiple strong variables are competing with each other to enter the model first. Using the max-$t$ $p$-values, ForwardStop selects a model of size 8, compared to size 3 when using the saturated-model $p$-values.

Because the null max-$t$ $p$-values are independent, while the saturated-model $p$-values are not, ForwardStop is guaranteed to control FDR when using the former but not the latter. In Section~\ref{sec:pValsIndep} we discuss intuitive sufficient conditions for independence and show that the max-$t$ and many other selected-model tests satisfy these conditions.

\subsection{Outline}

For the remainder of the article we will discuss the problem of selective sequential model selection in some generality, returning periodically to the selective max-$t$ test in forward stepwise regression and its lasso regression counterpart, the next-$\lambda$ test, as intuitive and practically important running examples.

In comparison to their saturated-model counterparts, we will see that the selected-model tests proposed here have three main advantages: power, independence (in many cases), and generalizability beyond linear regression. These advantages do not come entirely for free, as we will see: the selected-model methods test a more restrictive null hypothesis than the saturated-model methods. In addition, the selected-model tests require accept-reject or MCMC sampling, while the saturated-model tests can be carried out in closed form.

Section~\ref{sec:genericSetting} introduces the general problem setting, gives important definitions, and reviews relevant literature on testing ordered hypotheses using a sequence of $p$-values. Many of these methods require the $p$-values to be uniform and independent under the null, and we derive general conditions to ensure this property in Section~\ref{sec:pValsIndep}. In Section~\ref{sec:selective-reg} we contrast selected-model and saturated-model tests for linear regression setting, and illustrate why selected-model tests are often much more powerful in early steps of the path. In Section~\ref{sec:computation} we discuss strategies for sampling from the null distribution for our tests, and prove that in forward stepwise regression and $\ell_1$-regularized likelihood methods, the number of constraints in the sampling problem is never more than twice the number of variables. We provide a simulation study in Section~\ref{sec:sparseReg} and briefly discuss applications of our framework to decision trees, nonparametric changepoint detection, and principal components analysis in Section~\ref{sec:further-examples}. The paper ends with a discussion.

\section{Selective Sequential Model Selection}
\label{sec:genericSetting}

In the general setting, we observe data $Y \in \cY$ with unknown sampling distribution $F$, and then apply some algorithm to generate an adaptive sequence of $d$ nested statistical models contained in some {\em upper model} $M_{\infty}$.
\[
M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub M_\infty.
\]
By {\em model}, we mean a family of candidate probability distributions for $Y$. For example in  linear regression, a model specifies  the set of
predictors allowed to have nonzero coefficients (but not the  values of their coefficients). Without loss of generality, we can take $M_\infty$ as the union of all models under consideration. We will use the notation $M_{0:d}$ to denote the sequence $M_0, \ldots, M_d$, which we call the {\em model path}. We assume throughout that for each candidate model $M$ there is a minimal sufficient statistic $U(Y; \,M)$, and write
\[
U_k(Y) = U(Y; \,M_k(Y)).
\]

Given data $Y$ and model path $M_{0:d}$, we set ourselves the formal goal of selecting the simplest correct model: the smallest $k$ for which $F\in M_k(Y)$. Of course, in most real data problems all of the $M_k(Y)$ in the path, as well as all other models under consideration, are incorrect. In more informal terms, then, we seek the simplest model in the path that is not refuted by the available data.

Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model. By construction, $F\in M_k \iff k \geq k_0$. A {\em stopping rule} is any estimator $\hk(Y)$ of $k_0$, with $M_k$ ``rejected'' if $k < \hk$, and ``accepted'' otherwise. Because $\hk$ is the number of models we do reject, and $k_0$ is the number we should reject, the number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$. Depending on the scientific context, we might want to control the FWER $\P(V>0)$, the FDR $\E[V/\hk; \hk>0]$ , or another error rate defined by the expectation of some loss function $g(\hk, k_0)$:
\begin{equation}\label{eq:errRate}
\Err_F(\hk(\cdot), g) = \E_F\left[ g\left(\hk(Y), k_0(Y)\right)\right]
\end{equation}

\subsection{Single-Step $p$-Values}\label{sec:singleStep}

We will restrict our attention to stopping rules like those recently proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, which operate on a sequence $p_{1:d}$ of $p$-values. At each step $k$, we will construct a $p$-value for testing
\[
H_{k}:\; F\in M_{k-1}(Y)
\]
against the alternative that $F\in M_\infty\setminus M_{k-1}$, accounting for the fact that $M_{k-1}$ is chosen adaptively.

Following \citet{fithian2014optimal}, we say that for a fixed candidate model $M$, the random variable $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is conservative under sampling from any $F\in M$, given that $M$ is selected. That is,
\begin{equation}
\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
\end{equation}

Once we have constructed selective $p$-values for each fixed $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
\[
p_k(y) = p_{k, M_{k-1}(y)}(y),
\]
which is conservative on the event $\{F \in M_{k-1}(Y)\}$:
\begin{equation}\label{eq:selectiveGuaranteePk}
\P_F\left(p_k \leq \alpha \mid F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\end{equation}

One useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model $M_{k-1}$. By sufficiency, the law
\[
\L(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
\]
is the same for every $F\in M$. Thus, we can construct selective tests by comparing the realized value of any test statistic $T(Y)$ to its known conditional distribution under the null. All that remains is to choose the test statistic and compute its conditional null distribution, which can be challenging. See \citet{fithian2014optimal} for a general treatment.

\subsection{Sparse Parametric Models}\label{sec:genSparse}

Many familiar path algorithms including lasso, LARS, and forward stepwise regression, are methods for adaptively selecting a set of predictors in linear regression models where we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors. For each possible active set $E \sub \{1,\ldots,p\}$, there is a corresponding candidate model
\[
M(E):\; Y \sim \cN( X_E\beta_E, \sigma^2 I_n),
\]
which is a subset of the {\em full model}
\[
M_\infty:\; Y \sim \cN(X\beta, \sigma^2I_n).
\]
If the error variance $\sigma^2$ is known, the complete sufficient statistic for $M(E)$ is $X_E'Y$; otherwise it is $\left(X_E'Y,\, \|Y\|^2\right)$.

For the most part, we can discuss our theory and proposals in a generic parametric setting generalizing the linear regression problem above. Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^{\cJ}$:
\[
M_\infty = \{F_\theta:\; \theta \in \Theta\}.
\]
For any subset $E\sub \cJ$ define the sparse submodel with active set $E$ as follows:
\[
\Theta(E) = \{\theta:\; \theta_j = 0, \;\;\forall j \notin E\}, 
\quad M(E) = \{F_\theta:\; \theta\in \Theta(E)\}.
\]

We can consider path algorithms that return a sequence of nested active sets
\[
E_0(Y) \sub E_1(Y) \sub \cdots \sub E_d(Y) \sub \cJ,
\]
inducing a model path with $M_k = M(E_k),$ for $k=0,\ldots,d$. We will be especially interested in two generic path algorithms for the sparse parametric setting: forward stepwise paths and ever-active regularization paths.

\subsubsection{Forward Stepwise Paths and Greedy Likelihood Ratio Tests}
Let $\ell(\theta; Y)$ denote the log-likelihood for model
$M_\infty$. The {\em forward stepwise} algorithm proceeds as follows: we begin with some fixed $E_0$ (such as an intercept-only model), then at step $k=1,\ldots,d$, we set
\begin{equation}
j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
E_k = E_{k-1} \cup \{j_k\}.
\end{equation}
That is, at each step we select the next variable to maximize the likelihood of the next fitted model.

A natural choice of test statistic is the {\em greedy likelihood ratio statistic}
\begin{equation}\label{eq:greedyLRT}
G_k(Y) = \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)},
\end{equation}
the generalized likelihood ratio statistic for testing $M(E_{k-1})$ against the ``greedy'' alternative with one more active parameter:
\begin{equation}\label{eq:greedyAlternative}
M(E_{k-1})^{+1} = \bigcup_{j \notin E_{k-1}} M(E_{k-1} \cup \{j\}).
\end{equation}

The {\em selective greedy likelihood ratio test} rejects for large $G_k$, based on the law
\begin{equation}
\L\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
\end{equation}

Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood ratio test. The counterpart of the max-$t$ test in linear regression with known $\sigma^2$ is the max-$z$ test, which differs only in replacing the $t$-statistics $t_{j,E}$ with their corresponding $z$-statistics. The max-$z$ test is also equivalent to the selective greedy likelihood ratio test.

For simplicity, we have made two assumptions: that only one variable is added at each step, and that the set of candidate variables we choose from is the same in each step. It is relatively straightforward to relax either assumption, but we do not pursue such generalizations here.

\subsubsection{Ever-Active Regularization Paths and next-$\lambda$ Tests}
Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path. 
The notion of ever-active sets is needed since these solution paths can drop  (as well as add) predictors along the way. 	For some ordered set $\Lambda$, let $\lambda\in\Lambda$ parameterize the regularization penalty $P_\lambda(\theta)$. In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty]$.

\JTcomment{Yes, this looks a little funny to me too. Not all penalties produce
sparsity (for example, the generalized LASSO). So, I think this section is really thinking 
of $\ell_1$ norm (maybe a group LASSO norm). As written, 
we are also assuming that
there is a unique solution to (11) (which is not a big assumption but we might
as well make it.) Finally, $\ell$ does not actually have to be a likelihood.
BUT later on we implicitly assume that it is the likelihood for 
$M_{\infty}$. Why not just say this here, i.e. when talking
about $\ell_1$ regularization paths the models we have in mind are typically
going to be ones which correspond to a likelihood for some ``largest model'' for which we are trying to find a sub-model.

An example where we might not take $\ell$ to be a likelihood would be
if it were a pseudo-likelihood as in the neighborhood selection 
instead of graphical LASSO.
}

Assuming for simplicity that there is a unique solution to each penalized problem of the form
\begin{equation}\label{eq:pen-lik}
  \hat\theta^{\lambda}(Y) =
  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta),
\end{equation}
we define the ever-active set for $\lambda\in\Lambda$ as
\begin{equation}
  \tE_\lambda(Y) = \left\{j:\; \hat\theta_j^\nu(Y) \neq 0 
    \text{ for any } \nu\in \Lambda, \nu \geq \lambda \right\}
\end{equation}
It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many values.

Note that the ever-active sets $\tE_\lambda$ are nested by construction. In addition, we will assume that $|\tE_\lambda|<\infty$ for every $\lambda\in\Lambda$. Our model path will correspond to the sequence of distinct ever-active sets. 
Formally, let 
\[
\lambda_0=\sup \Lambda, \quad \text{ and } \quad 
E_0 = \bigcap_{\lambda\in \Lambda} \tE_\lambda,
\]
and for $k\geq 1$ let $\lambda_k$ denote the (random) value of $\lambda$ where the active set changes for the $k$th time:
\begin{equation}
  \Lambda_k = \{\lambda\in \Lambda:\; \tE_\lambda \supsetneq \tE_{\lambda_{k-1}}\},
  \quad
  \lambda_k = \sup\Lambda_k,
  \quad \text{ and } \quad
  E_k = \bigcap_{\lambda\in\Lambda_k} \tE_{\lambda}.
\end{equation}

In this setting, $\lambda_k$ is a natural test statistic for model $M_{k-1}$, with larger values suggesting a poorer fit. The {\em selective next-$\lambda$ test} is the test that rejects for large $\lambda_k$, based on the law
\begin{equation}
\L\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
\end{equation}

\JTcomment{The next-$\lambda$ test is actually a little expensive to compute if we a finite set of $\lambda$  as, say, glmnet does. Should we not mention this? On the plus side, 
it gives us tests that do not require LAR and can be used with glmnet. 
Formally, to carry this out test, we need to draw samples of $Y$ conditioned
on the necessary sufficient statistics satisfying the necessary constraints 
(for LASSO, there $2(p-k)$ constraints on $X_{-E}^Ty$). Then, for each sample
we restart glmnet and run until the active set changes, storing this
value of $\lambda$. Then we compare our observed next-$\lambda$ to this
sample of $\lambda$'s.

In the LAR-LASSO case, this is easier as there is an explicit formula
to compute the next event. But, LAR is limited to LASSO whereas next-$\lambda$
in principle works for group LASSO.

Finally, the simplification of constraints we get really only works for
almost only for $\ell_1$. For group LASSO, the simplification
doesn't happen. The constraints after $k$ steps look like
$$
\left\{y: P_Ey=z, \|X_g^T(y - \hat{\mu}_l(z))\|_2 \leq \lambda_l \omega_g, 
g \in -E, 1 \leq l \leq k, \|(I-P_E)y\|=c\right\}
$$
where $z$ is the projection of our observed $y$ onto the ever active set.
So, for a given ``ever inactive group'', this says that $X_g^Ty$ is in
an intersection of $k$ balls which does not necessarily have a simple structure.
When all the groups are of size 1, i.e. in the LASSO, the intersection 
of $k$ balls is the intersection of $k$ intervals and we get a simplification.

All this to say, that we are really only talking about $\ell_1$ here, and 
next-$\lambda$ is essentially only practical BLAH
}

\paragraph{Remark} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion:
\[
\hat\beta_{\text{lasso}}^\lambda = \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1.
\]

This is slightly different from the penalized log-likelihood criterion of \eqref{eq:pen-lik}, which would minimize
\begin{equation}\label{eq:lasso-beta-sigma}
  -\ell(\beta, \sigma;\,Y) + P_\lambda(\beta)=
  \frac{n}{2}\log(2\pi\sigma^2) + \frac{1}{2\sigma^2}\|Y-X\beta\|_2^2 
  + \lambda\|\beta\|_1.
\end{equation}

If $\sigma^2$ is known, the two methods are equivalent up to rescaling $\lambda$. If $\sigma^2$ is unknown, then partially minimizing~\eqref{eq:lasso-beta-sigma} over $\sigma^2$ gives $\hat\sigma^2(\beta) = n^{-1}\|Y-X\beta\|_2^2$, leading to profile log-likelihood
\[
-\ell^*(\beta;\,Y) \;\triangleq\; -\ell(\beta, \hat\sigma^2(\beta); \,Y)
\;=\;  \frac{n}{2}
\left(\log\left(\frac{2\pi\|Y-X\beta\|_2^2}{n}\right)+1\right)
\]

As a result the lasso coefficient vector $\hat\beta_{\text{lasso}}^\lambda$ minimizes
\[
\frac{n}{2\pi e}\exp\left\{-2\ell^*(\beta;\,Y)/n\right\} + \lambda\|\beta\|_1,
\]
while the regularized log-likelihood estimator $\hat\beta^\lambda$ minimizes $-\ell^*(\beta;\,Y)+\lambda\|\beta\|_1$. If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path. Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and the theory in this article would go through unchanged.

\begin{comment}
\begin{align*}
  \hat\sigma^2(\beta)
  &\triangleq \argmax_{\sigma^2>0} \ell(\beta,\sigma^2;\,Y) 
  = \frac{1}{n}\|Y-X\beta\|_2^2, \quad \text{ and }
  \\
  \ell^*(\beta; \,Y)
  &\triangleq \sup_{\sigma^2>0} \ell(\beta, \sigma^2;\,Y) 
  = -\frac{n}{2} - \frac{n}{2}\log\left(\frac{2\pi\|Y-X\beta\|_2^2}{n}\right).
\end{align*}

As a result the lasso coefficients $\hat\beta_{\text{lasso}}^\lambda$ minimize
\[
\exp\{-2\ell(\beta,\sigma;\,Y)/n \,+\, \text{const.}\} + \lambda\|\beta\|_1,
\]
instead of the criterion in \eqref{eq:pen-lik}. If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path. Otherwise, we could ``fix'' this issue (at the cost of some notational bloat) by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and all of the theory in this article would go through unchanged.
\end{comment}

\subsection{Which Null Hypothesis?}
\label{sec:whichnull}

Note that in our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$. Even if the next selected variable $X_{j_k}$ is a noise variable in the full model, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are other signal variables that have not yet been included.

In some circumstances, we might want to define a type I error at step $k$ differently, by choosing a different null hypothesis to test. Let $\mu = \E Y$ and let $\theta^E$ denote the {\em least-squares coefficients} of active set $E$ --- the coefficients of the best linear predictor for the design matrix $X_E$:
\[
\theta^E = X_E^+ \mu = \argmin_{\theta\in \R^{|E|}} \|\mu - X_E\theta\|_2^2,
\]
where $X_E^+$ is the Moore-Penrose pseudoinverse of the matrix $X_E$. 
These are the true ``partial regression coefficients''  for design matrix $X_E$.
 Write $\proj_E\mu = X_E \theta^E$ for the projection of $\mu$ into the column space of $X_E$.

\citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression:
\begin{description}
\item[Complete Null:] $M_{k-1}$ is (already) correct. That is, 
\[
H_k:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}.
\]
\item[Incremental Null:] $M_{k-1}$ may be incorrect, but $M_k$ is no improvement. That is, 
\[
H_k^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0.
\]
\item[Full-Model Null:] The coefficient of $X_{j_k}$ is zero in the ``full'' model with all $p$ predictors. That is,
\[
H_k^{\text{full}}:\; \theta_{j_k}^{\{1,\ldots,p\}} = 0.
\]
\end{description}

While the complete null is the strongest hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. Defining 
\begin{align}
V^{\text{inc}} &= \#\{k < \hk:\; H_k^{\text{inc}} \text{ is true}\} \quad \text{ and } \\
V^{\text{full}} &= \#\{k < \hk:\; H_k^{\text{full}} \text{ is true}\},
\end{align}
we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates. For example, we could define
\[
\text{FDR}^{\text{full}} = \E[V^{\text{full}} / (\hk \vee 1)],
\]
as the false discovery rate with respect to the full-model null. \citet{barber2014controlling} present a framework for controlling $\text{FDR}^{\text{full}}$.

The full-model null is the most different conceptually from the other two, taking a ``variable-centric'' instead of ``model-centric'' viewpoint, where it is of primary scientific interest to discover variables that have nonzero coefficients after controlling for all $p-1$ other variables that were measured.

To elucidate this distinction consider two methods we could apply in a classical setting where we have some fixed {\em a priori} ordering for the variables, inducing the fixed sequence of nested linear models
\[
M(\emptyset) \sub M(\{1\}) \sub M(\{1,2\}) \sub \cdots \sub M(\{1,\ldots, p\}).
\]

In the first procedure, we construct a (type I) analysis-of-variance (ANOVA) table performing an $F$- or $t$-test of each model against the next. We then stop and select model $M_k=M(\{1,\ldots,k\})$ the first time the $F$-test $p$-value is larger than $\alpha$. This procedure would control the complete-null and incremental-null FWER, but {\em not} the full-model-null FWER. The methods we consider in this article are essentially generalizations of this procedure for the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.

In the second procedure, we fit the full model $M(\{1,\ldots,p\})$ and compute a $t$-statistic for each variable with respect to that model, yielding $t$-statistics $t_1, t_2, \ldots, t_p$ (not ordered by size). For $k=1,\ldots,p$, we compare each $|t_j|$ to the level-$\alpha$ cutoff for a $t$-test with $n-p$ degrees of freedom, and stop selecting variables the first time $|t_{k+1}|$ is not larger than this cutoff. We might use this procedure if the variables had a  fixed ordering.
This procedure would control the full-model-null FWER, but {\em not} the incremental-null FWER. Note that the {\em model} $M_k$ plays no role in this procedure: its fit to the data is never evaluated and it is not really selected in any meaningful sense. Even if the fit of $M_k$ is far worse than that of $M_{k+1}$, we may not be able to select $X_{k+1}$ if (say) it is highly collinear with $X_{k+100}$. Conversely, even if the fit of $M_k$ is no better than that of $M_{k-1}$, we may still select $X_k$ if it is a highly useful predictor when used in concert with $X_{k+100}$.

More generally, methods that control full-model error rates are best viewed not as {\em model-selection} procedures --- since all inferences are made with respect to the full model, which is specified ahead of time --- but {\em variable-selection} procedures that test multiple hypotheses with respect to a single fixed model. Note that the truth or falsehood of these hypotheses can depend sensitively on the set of $p$ predictors: rejecting $H_j^{\text{full}}$ has no meaning without reference to the list of all variables that we controlled for, which can make a rejection difficult to interpret when $p$ is large. Thus, error rates like $\text{FDR}^{\text{full}}$ are best motivated when the full model has some special scientific status. For example, the scientist may believe due to theoretical considerations that the linear model in $X_1,\ldots,X_p$ is fairly credible, and that a nonzero coefficient $\beta_j$, controlling for all of the other variables, would constitute evidence for a causal effect of $X_j$ on the response.

In this article we will concern ourselves primarily with testing the complete null, reflecting our stated aim of choosing the least complex model that is not rejected by the data. As we discuss further in Section~\ref{sec:selective-reg}, the saturated-model tests of \citet{taylor2014exact} are valid selective tests of $H_k^{\text{inc}}$. Unfortunately, this requires us to assume $\sigma^2$ is known, can result in a large reduction in power, creates dependence between $p$-values at different steps, and is difficult to generalize beyond the case of linear regression. See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between the three approaches.

\subsection{Related Work}

The problem of model selection is an old one, with quite an extensive literature. However, with the exception of the works above, very few methods offer finite-sample guarantees on the model that is selected except in the orthogonal-design case. One  exception is the knockoff filter of \citet{barber2014controlling}, a variant of which provably controls the full-model FDR.
We compare our proposal to the knockoff method in Section \ref{sec:sparseReg}.

Methods like AIC \citet{akaike1974new} and BIC \citep{schwarz1978estimating} are designed for the non-adaptive case, where the sequence of models is determined in advance of observing the data. Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse (e.g. \cite{LY2015}).   \citet{benjamini2009simple} extend the AIC by using an adaptive penalty $\sigma^2\lambda_{k,p} k$ to select a model, based on generalizing the Benjamini-Hochberg procedure \citet{benjamini1995controlling}, but do not prove finite-sample control. The stability selection approach of \citet{meinshausen2010stability} uses an approach based on splitting the data many times and offers asymptotic control of the FDR, but no finite-sample guarantees are available.

\section{Stopping Rules, Ordered Testing, and Independence}

An ordered hypothesis testing procedure takes in a sequence of $p$-values $p_1, \ldots, p_d$ for null hypotheses $H_{1}, \ldots, H_{d}$, and outputs a decision $\hk$ to reject the initial block $H_{1}, \ldots, H_{\hk}$ and accept the remaining hypotheses. Note that in our setup the hypotheses are nested, with $H_{k-1} \sub H_k$; as a result, all of the false hypotheses precede all of the true ones.

We first review several proposals for ordered-testing procedures, several of which require independence of the null $p$-values conditional on the non-null ones. These procedures also assume the sequence of hypotheses is fixed, whereas in our setting the truth or falsehood of $H_k$ is random, depending on which model is selected at step $k$. However, we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on $k_0$.

\subsection{Proposals for Ordered Testing of Fixed Hypotheses}
\label{sec:orderedProposals}

We now review several proposals for ordered testing of fixed null hypotheses along with their error guarantees in the case where the sequence of null hypotheses is fixed. In the next section we will extend the analysis to random null hypotheses.

\subsubsection{BasicStop}
The simplest procedure is to keep rejecting until the first time that $p_k > \alpha$: 
\[
\hk_B(Y) = \min\left\{k:\; p_k > \alpha\right\} - 1
\]
We will call this procedure {\em BasicStop}. It is discussed in \citet{marcus1976}.

Since $k_0$ is the index of the first null hypothesis, we have
\begin{equation}\label{eq:basic-stop-control}
 \text{FWER} \leq \P(p_{k_0} \leq \alpha) \leq alpha.
\end{equation}

\subsubsection{ForwardStop}

\citet{gsell2013sequential} propose the {\em ForwardStop} rule to controls the FDR:
\[
  \hk_{F}(Y) = \max\left\{k:\;
    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
\]
ForwardStop uses the $p$-values {\it before} step $k$ in making its decision whether to reject at step $k$. If $p_k$ is uniform, then $-\log(1-p_k)$ is an $\text{Exp}(1)$ random variable with expectation 1; thus, the sum can be seen as an estimate of the false discovery proportion (FDP):
\[
\widehat{\text{FDP}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i).
\]
ForwardStop chooses the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$. \citet{gsell2013sequential} show that ForwardStop controls the FDR if the null $p$-values are independent of each other and of the non-nulls.

\subsubsection{Accumulation Tests and HingeExp}

\citet{li2015accumulation} generalize ForwardStop, introducing the family of {\em accumulation tests}, defined as follows.
Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int_{t=0}^1 h(t)dt=1$, called an {\em accumulation function.} Using $h$ we can construct an estimate of FDP at step $k$:
\begin{equation}
\widehat{\text{FDP}}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k},
\end{equation}
and select the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$. Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
\begin{equation}
  h_{\rm HingeExp}(p)= 
  \begin{cases} 
    C\log\left(\frac{1}{C(1-p)} \right)& p> 1-1/C\\
    0 & p\leq 1-1/C
  \end{cases}
\end{equation}

\citet{li2015accumulation} show that accumulation tests control a modified FDR criterion provided that the null $p$-values are $U[0,1]$, and are independent of each other and of the non-nulls.

\subsection{Ordered Testing of Random Hypotheses}

In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section. It is not immediately clear, then, how much we should condition on when we construct the $p$-value at step $k$. In particular, is it sufficient to condition only on the current null model $M_{k-1}$, or must we condition on the entire subpath of models $M_0,\ldots,M_{k-1}$?

Note that conditioning on the current null model {\em is} enough to guarantee that $p_k$ is uniform conditional on the event $\{F\in M_{k-1}(Y)\}$ that the $k$th null is true. Unfortunately, however, it is {\em not} enough to guarantee that the above stopping rules actually control FDR or FWER.

Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution and consider the following counterexample:

\begin{restatable}{proposition}{counterex}\label{prop:counterexample}
  Consider linear regression with $n=p=3$, known $\sigma^2=1$ and identity design $X=I_3$.

  Suppose that we construct the model path as follows: if $|Y_3|>z_{\alpha/2}$, we add $X_1$ to the active set first, then $X_2$, then $X_3$. Otherwise, we add $X_2$, then $X_1$, then $X_3$. At each step we construct selective max-$z$ $p$-values and finally choose $\hat k$ using BasicStop.

  If $\mu=(0,C,0)$, then the FWER for the entire procedure tends to $2\alpha-\alpha^2$ as $C\to\infty$.
\end{restatable}

A proof of Proposition~\ref{prop:counterexample} is given in the appendix. The crucial issue is that $k_0$ is no longer a fixed index. Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value $p_{k_0}$ corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.

If, however, we could guarantee that
\[
\P(p_k \leq \alpha \mid H_k \text{ true}, k_0=k) \leq \alpha
\]
then BasicStop again would control FWER, by the argument in \ref{eq:basic-stop-control}.

Note that $k_0$ is unknown, so we cannot directly condition on its value when we construct our single-step $p$-values. However, because
\[
\{k_0=k\} = \{F\in M_{k-1}, F\notin M_{k-2}\},
\]
it is enough to condition on $(M_{k-2}, M_{k-1})$ when we construct our $p$-values. As we will see in Section~\ref{sec:pValsIndep}, conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{0:k-1}, U_{k-1})$ in most cases of interest including forward stepwise likelihood paths and ever-active regularization paths (but not the path in Proposition~\ref{prop:counterexample}).

Similarly, the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply to case where the null hypotheses are random. However, we recover these guarantees if we have {\em conditional} independence and uniformity of null $p$-values given $k_0$:
\begin{equation}\label{eq:indepCond_random_k0}
  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_k, \; k_0 = k) = \prod_{i=k+1}^d \alpha_i
\end{equation}
For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.

The following proposition shows that independence on nulls allows us to transfer the error-control guarantees of ForwardStop and accumulation tests to the random-hypothesis case.
\begin{proposition}
  Let $\hk$ be a stopping rule operating on $p$-values $p_{1:d}(Y)$ for nested hypotheses $H_{1:d}(Y)$. For some function $g$, let
  \[
  \Err_g = \E\left[ g\left(\hk(p_{1:d}(Y)), k_0(Y)\right)\right],
  \]

  Suppose that $\hk$ controls $\Err_g$ at level $\alpha$ if the $H_{1:d}$ are fixed and the null $p$-values are uniform and independent of each other and the non-nulls. Then, $\hk$ controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
\end{proposition}
\begin{proof}
    For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If $\hk$ controls $\Err_g$ in the 
fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$. Thus, \eqref{eq:indepCond_random_k0} implies
  \[
  \E\left[ g\left(\hk(p_1,\ldots,p_d), k_0(Y)\right) \;\mid\;
      k_0(Y), \; p_{1:k_0}(Y)\right] \leqAS \alpha.
  \]
  Marginalizing over $(p_{1:k_0},k_0)$ gives $\Err_g\leq \alpha$.
\end{proof}

Note that~\eqref{eq:indepCond_random_k0} implies in particular that each $p_k$ is uniform given $k_0=k$; thus, independence on nulls is enough to guarantee that BasicStop and ForwardStop control FWER and FDR, respectively. The next section discusses conditions on the model sequence $M_{0:d}$ and the $p$-value sequence $p_{1:d}$ under which~\eqref{eq:indepCond_random_k0} is satisfied.

\section{Conditions for Independent $p$-values}\label{sec:pValsIndep}

We now develop sufficient conditions for constructing $p$-values with the {\it independent on nulls} property (\ref{eq:indepCond_random_k0}). To begin, we motivate the general theory by discussing a specific case, the max-$t$ test for forward stepwise regression.

\subsection{Independence of max-$t$ $p$-values}

Recall that at step $k$, the max-$t$ test rejects for large $T_k=|t_{E_{k-1}}^*|$, comparing its distribution to the conditional law
\begin{equation}\label{eq:maxTnull}
\L\left(T_k \; \mid \; E_{k-1}, \,X_{E_{k-1}}'Y, \,\|Y\|^2 \right).
\end{equation}
This conditional distribution is the same for all $F\in M(E_{k-1})$, because $U_{k-1}=(X_{E_{k-1}}'Y,\; \|Y\|^2)$ is the complete sufficient statistic for $M(E_{k-1})$.

Now, if $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:k-1}$, for two reasons:
\begin{enumerate}
\item $p_k$ is uniform conditional on $E_{k-1}$ and $U_{k-1}$ by construction, and
\item $p_1, \ldots, p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$, as we will see shortly.
\end{enumerate}

Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:k-1}$, guaranteeing that $\L(p_k \mid p_{1:k-1}) = U[0,1]$ whenever $H_k$ is true.

Next we will explain why $p_1,\ldots,p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$. Knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing $(X_{E_{k-1}}'Y,\|Y\|^2)$ is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$. As a result, we can reconstruct the order in which those $k-1$ variables were added, since for $s<k$, 
\[
j_s = \argmax_{j \notin E_{s-1}} |t_{j,E_{s-1}}| 
= \argmax_{j \in E\setminus E_{s-1}} |t_{j,E_{s-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
\]
In general, we say that the path $M_{0:d}$ follows the {\em subpath sufficiency principle} (SSP) if for each $k$ the subpath $M_{0:k}$ is a function of $(M_k,U_k)$.

Furthermore, for $s<k$, $p_s$ is computed by comparing $T_s=|t_{j_s, E_{s-1}}|$ to the upper $\alpha$ quantile of the distribution $\L(T_s \mid E_{s-1}, U_{s-1})$, computed under the assumption $F\in M(E_{s-1})$. The conditional null distribution is a function of $(E_{s-1}, U_{s-1})$, and as we have already seen, the test statistic $T_s$ is a function of $(E_s, U_s)$, both of which are functions of $(E_{k-1},U_{k-1})$.

Because $(E_s,U_s)$ is a function of $(E_k,U_k)$ for $s<k$, we can characterize the above discussion in terms of the filtration $\sF_{0:d}$, where
\[
\sF_k = \sF(E_k, U_k),
\]
the $\sigma$-algebra generated by random variables $(E_k, U_k)$. We could equivalently write $\sF_k=\sF(M_k, U_k)$, or $\sF_k=\sF(M_{0:k}, U_k)$.

At each $k$, given that $H_k$ is true, $\L(p_k \mid \sF_{k-1})\sim U[0,1]$, and $p_k$ is $\sF_k$-measurable. If $H_k$ is true, then so are $H_{k+1},\ldots, H_d$, and applying this conclusion iteratively we can see that all of the remaining $p$-values are uniform and independent.

By contrast, the saturated model $p$-values use a law different from~\eqref{eq:maxTnull} to compute $p$-values, one that uses
information not contained in $(E_k,U_k)$. As a result, saturated-model $p$-values are generally not independent on nulls. We discuss the regression setting in more detail in Section \ref{sec:selective-reg}.

\subsection{General Case}

We now present sufficient conditions for independence on nulls generalizing the above ideas, culminating in Theorem~\ref{thm:suffCond} at the end of this section.

A {\em selective $p$-value} $p_k$ for testing $H_k:\; F \in M_{k-1}(Y)$ satisfies for any $F\in M_{\infty}$,
\[
\P_F(p_k \leq \alpha \mid M_{k-1},\; F\in M_{k-1}) \leqAS \alpha.
\]

We say that a filtration $\sF_{0:d}$ {\em separates} the $p$-values $p_{1:d}$ if
\begin{enumerate}
\item $M_k(Y)$ and $p_k(Y)$ are measurable with respect to $\sF_k$, and 
\item $p_k(Y)$ is uniform or conservative given $\sF_{k-1}$ on the event $\{F\in M_{k-1}(Y)\}$.
\end{enumerate}
Separated $p$-values are independent on nulls, as we see next.

\begin{proposition}[Independence of 
  Separated $p$-Values]\label{prop:jointConserv}

  Let $p_{1:d}$ be selective $p$-values for $M_{0:d}$, 
  separated by $\sF_{0:d}$.

  If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
  independent and uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
  If they are conservative, then for all
  $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
  \begin{equation}\label{eq:jointConserv}
  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d\;
    \mid\; k_0 \leq k, \; \sF_k\right) \leqAS \prod_{i=k+1}^d
  \alpha_i.
  \end{equation}
\end{proposition}

\begin{proof}
  We first prove~\eqref{eq:jointConserv}
  by induction. Define $B_i = \{p_i \leq \alpha_i\}$. 
  The base case is
  \[
  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d,
  \]
  which is true by construction of $p_d$. 
  For the inductive case, note that the
  conditional probability on the left-hand side
  of~\eqref{eq:jointConserv} is
  \begin{align*}
    \P_F\left(B_{k+1}, \ldots, B_d
      \mid \sF_k\right)1_{\{k_0 \leq k\}} 
    &\eqAS \E_F\bigg[ 1_{B_{k+1}} 
    \P_F\left(B_{k+2}, \ldots, B_d
      \mid \sF_{k+1}\right)1_{\{k_0 \leq k+1\}}
    \mid \sF_k\bigg]1_{\{k_0 \leq k\}}\\
    &\leqAS \P_F\left[ B_{k+1}
      \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
    \quad\;\;\leqAS\;\; \prod_{i=k+1}^d \alpha_i
  \end{align*}

  If the $p_k$ are exact then the above 
  inequalities become equalities, implying uniformity and joint independence.
\end{proof}

If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.

Define the {\em sufficient filtration} for the path $M_{0:d}$ as the filtration $\sF_{0:d}$ with $\sF_k = \sF(M_{0:k}, U_k)$. By our assumption of minimal sufficiency, $U_s$ is always a function of $(M_s, U_k)$ for $s<k$. The path algorithm satisfies the SSP if and only if
\begin{equation}\label{eq:SSP}
  \sF(M_{0:k}, U_k) = \sF(M_k, U_k), \quad k=0,\ldots, d
\end{equation}

The sufficient filtration separates $p_{1:d}$ if and only if 
\begin{enumerate}
\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and \RTcomment{clearer to say ``sub-uniform''?} \JTcomment{I think you mean ``super-uniform'', i.e. it is larger than a uniform random variable?}
\item $p_k$ is a function of $M_{0:k}$ and $U_k$.
\end{enumerate}

To be a valid selective $p$-value for a single step, $p_k$ only needs to be conservative conditional on $M_{k-1}$. In general, it is more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:k-1}$ as well as $U_{k-1}$. In most cases of interest, however, (1) we will condition on $(M_{k-1}, U_{k-1})$ anyway, to eliminate nuisance parameters, and (2) the path satisifies the SSP, so conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$. Thus, most single-step selective $p$-values we would contemplate using are already exact conditional on $\sF_{k-1}$.

The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite. For example, we will see in Section~\ref{sec:selective-reg} that it excludes saturated-model tests in linear regression.

Collecting together the threads of this section, we arrive at our main result: a general sufficient condition for a $p$-value sequence to be independent on nulls.
\begin{theorem}[Sufficient Condition for Independence on Nulls] \label{thm:suffCond}
Assume that $U(Y; M)$ is a complete sufficient statistic for each candidate model $M$, that each $p_k$ is an exact selective $p$-value for $M_{k-1}$, and that the path algorithm $M_{0:d}(\cdot)$ satisfies the SSP.

If $p_k$ is $\sF_k$-measurable for each $k$, then the $p$-value sequence $p_{1:d}(Y)$ is independent on nulls.
\end{theorem}

\begin{proof}
We apply the definition of completeness to the function 
\[
h(U(Y; \, M)) \;=\; 
\alpha - \P\bigg(p_k(Y) \leq \alpha \mid U(Y; \, M), \, M_{k-1}(Y) = M\bigg)
\]
If $p_k$ is exact given $M_{k-1} = M$, then 
\[
\E_F[h(U_{k-1}) \mid M_{k-1}=M] \;=\; 0
\] 
for every $F\in M$. Therefore, we must have $h(U_{k-1})\eqAS 0$, so $p_k$ is independent of $\sF(M_k, U_{k-1})$. Because $M_{0:d}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{0:k-1}, U_{k-1})$.

If $p_k$ is also $\sF_k$-measurable, then the sufficient filtration separates $p_{1:d}$, implying that the sequence $p_{1:d}$ is independent on nulls.
\end{proof}

Note that if our path algorithm did not satisfy the SSP, we could in many cases still repair the situation by constructing $p$-values that are uniform conditional on $(M_{0:k-1},U_{k-1})$, not just conditional on $(M_{k-1},U_{k-1})$.

\subsection{Independence for Greedy Likelihood and next-$\lambda$ $p$-values}\label{sec:SSP}

In this section we establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP. What is more, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration, and the next-$\lambda$ statistic $\lambda_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration. As a result, the selective greedy likelihood $p$-values in the forward stepwise path are independent on nulls, as are the next-$\lambda$ $p$-values in the regularized-likelihood path, per Theorem~\ref{thm:suffCond}.

Define the restricted log-likelihood
\begin{equation}
\ell_E(\theta; Y) = \left\{\begin{matrix} 
    \ell(\theta; \;Y) & \theta \in \Theta(E)\\ 
    -\infty  & \mathrm{ otherwise.}\end{matrix}\right. 
\end{equation}

\begin{proposition}\label{prop:forwardSSP}
  The forward stepwise path satisfies the subpath sufficiency principle.
\end{proposition}
\begin{proof}
  Recall the forward stepwise path is defined by
  \begin{equation}\label{eq:forwardStepDef}
    j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
    E_k = E_{k-1} \cup \{j_k\}
  \end{equation}

  For some fixed step $k$ and active set $E$, let $A$ denote the event $\{M_k = M(E)\}$. Conditioning on $U$, the restricted likelihood is proportional to a function depending only on $U = U(Y; \;M(E))$. The log-likelihood decomposes as
  \[
  \ell_E(\theta; Y) = \ell_E^{U}(\theta; U) 
  + \ell_E^{Y \mid U}(Y \mid U).
  \]
  The second term, the log-likelihood of $Y$ given $U$, does not depend on $\theta$ because $U$ is sufficient for $M(E)$.

  On $A$, we have $E_s(Y) \sub E$ for all $s\leq k$, meaning that the maximum in~\eqref{eq:forwardStepDef} is attained by some $j\in E$ at every step. So, for $s \leq k$, we have
\begin{align}
  j_s &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
    \theta\in\Theta(E_{s-1} \cup \{j\})\right\} \\
  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^U(\theta; U(Y)):\;
    \theta\in\Theta(E_{s-1} \cup \{j\})\right\}. \label{eq:onlyU}
\end{align}

Equation~\eqref{eq:onlyU} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $U(Y)$, which equals $U_k(Y)$ on $A$. As a result, it also follows that the entire sequence $M_{0:k}(Y)$ depends only on $(M_k, U_k)$.
\end{proof}

\JTcomment{Are we assuming that $U(Y;M)$ is minimal sufficient but not 
necessarily minimal
complete sufficient in the above proposition? The statement doesn't say
what $U$ satisfies. I ask because in the corollary we do say something
about completeness.  Our earlier assumption was just that $U$ always
denotes a minimal sufficient statistic.
}

\begin{corollary}\label{cor:greedyLikIndep}
  If $U(Y; M)$ is complete sufficient for model $M$, the selective greedy likelihood $p$-values computed for the forward stepwise path are independent on nulls.
\end{corollary}
\begin{proof}
  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration. First, the test statistic $G_k$ is $\sF_k$-measurable because
  \begin{align*}
    G_k(Y) &= \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)}\\
    &= \sup_{\theta\in \Theta(E_k)} {\ell_{E_k}^{Y\mid U}(\theta; U_k)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell_{E_k}^{Y\mid U}(\theta; U_k)}.
  \end{align*}
  Second, the null distribution $\L\left(G_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $G_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
\end{proof}

Analogous results hold for ever-active regularized likelihood paths and the next-$\lambda$ $p$-values.

\begin{proposition}\label{prop:regPathSSP}
Ever-active regularized likelihood paths satisfy the
subpath sufficiency principle.
\end{proposition}

\begin{proof}
  Again, for fixed $k$ and $E$ let $A$ denote the event $\{M_k = M(E)\}$, and define $U=U(Y;\, M(E))$. If we define
\[
\hat\theta^{(E,\lambda)} = \argmin_{\theta\in\Theta(E)} -\ell_E^{U}(\theta; U(Y)) + P_\lambda(\theta),
\]
then 
\[
\hat\theta^{(E,\lambda)} \eqAS \hat\theta^{\lambda} \text{ on } 
A \cap 1_{\{\lambda \geq \lambda_k\}}.
\]
But $\hat\theta^{(E,\lambda)}$ depends only on $U(Y)$. Therefore, on $A$, we can reconstruct the entire path of solutions for $\{\lambda\in \Lambda:\; \lambda\geq \lambda_k\}$, once we know $(M_k,\,U_k)$.
\end{proof}

\begin{corollary}\label{cor:nextLambdaIndep}
  If $U(Y; M)$ is complete sufficient for model $M$, the selective next-$\lambda$ $p$-values computed for any ever-active regularized likelihood path are independent on nulls.
\end{corollary}
\begin{proof}
  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.

  Recall that 
  \[
  \Lambda_k = \{\lambda\in \Lambda:\; \tE_\lambda \supsetneq \tE_{\lambda_{k-1}}\},
  \quad 
  \lambda_k = \sup\Lambda_k,
  \quad \text{ and } \quad
  E_k = \bigcap_{\lambda\in\Lambda_k} \tE_{\lambda},
  \]
  and that the $\tE_{\lambda}$ are nested by construction and finite by assumption. As a result, there is some $\nu_k\in \Lambda_k$ for which $\tE_{\nu_k} = E_k$.

  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \nu_k$, we have $\tE_\lambda \sub E_k$. As a result, ${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$ for such $\lambda$, so $\tE_{\lambda}$ depends only on $\ell_{E_k}^{Y\mid U}(Y;\, U_k)$, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.

  Second, the null distribution $\L\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
\end{proof}

As we have shown, the selective max-$t$ test is equivalent to the selective greedy likelihood test in the case of linear regression. In the next section, we will compare and contrast the max-$t$ and next-$\lambda$, and other {\em selected-model tests}, with the {\em saturated-model tests} proposed by \citet{taylor2014exact} and others.

\section{Selective $p$-Values in Regression}
\label{sec:selective-reg}
In linear regression, \citet{fithian2014optimal} draw a distinction between two main types of selective test that we might perform at step $k$: tests in the selected model, and tests in the saturated model. For simplicity, we will assume throughout this section that our path algorithm adds exactly one variable at each step. We also assume the path algorithm satisfies the SSP, so that we need not worry about the distinction between conditioning on $(M_{k-1}, U_{k-1})$ and conditioning on $(M_{0:k-1}, U_{k-1})$.

\subsection{Saturated-Model Tests}
Many conditional selective tests for linear regression use the saturated-model framework. See for example \citet{lockhart2014significance}, \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}. Define the saturated model as
\[
M_{\text{sat}}:\; Y \sim \cN(\mu, \sigma^2I_n),
\]
so named because there is a mean parameter for every observation. Because the parameter $\mu$ has the same dimension as the data $Y$, we must assume that $\sigma^2$ is known. If this is not the case, we can estimate $\sigma^2$ and apply the test while plugging in the estimated value; this is what we have done with the diabetes data to compute the $p$-values in Table~\ref{tab:diab}.

Saturated-model tests assume only that $M_{\text{sat}}$ is true, and perform inference on selected linear functionals of $\mu$. For example, in linear regression after selecting active set $E$, we could construct confidence intervals for the least-squares coefficients $\theta^E$ of active set $E$, defined as
\[
\theta^E \;\;=\;\; \argmin_{\theta\in\R^{|E|}} \|\mu - X_E\theta\| \;\;=\;\; (X_E'X_E)^{-1}X_E'\mu
\]

In the context of sequential model selection, we can use the saturated-model framework to test the incremental null 
\[
H^{\text{inc}}:\; \theta_{j_k}^{E_{k-1}\cup \{j_k\}} = 0.
\]
Let $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, and let $\proj_{\eta_k}^\perp$ denote the projection operator into the space orthogonal to $\eta_k$. The saturated model has $n$-dimensional natural parameter $\mu$ and complete sufficient statistic $Y$, and the UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
\begin{equation}\label{eq:satModel}
\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right),
\end{equation}
or equivalently on
\begin{equation}\label{eq:satModelEquiv}
\L\left( X_{j_k}'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right).
\end{equation}
The pair $(E_{k-1}, j_k)$ defines the selection event and $\proj_{\eta_k}^\perp Y$ eliminates the nuisance parameter $\proj_{\eta_k}^\perp \mu$.

While the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, the null distribution we compare it against is not: it depends on $\proj_{\eta_k}^\perp Y$, which is not measurable with respect to $\sF_k$. As a result, saturated-model $p$-values constructed as in~\eqref{eq:satModel} are in general {\em not} independent on nulls.

\subsection{Selected-Model Tests}
Selected-model inference~\citep{fithian2014optimal} 
represents a more powerful option when testing for inclusion of variable $j_k$. While saturated-model inference always uses the same model $M_{\text{sat}}$, selected-model inference uses the lower-dimensional statistical model that is chosen by our selection procedure, similarly to data splitting. In the sequential case this means that at step $k$, we test the selected null model $M(E_{k-1})$, either against the next selected model $M(E_k)\setminus M(E_{k-1})$ or against the upper model $M_\infty\setminus M(E_{k-1})$.

\subsubsection{Testing Against the Next Model}\label{sec:identify}
One option for selected-model inference is to perform a selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$, conditioning on both models. Conditioning on $(M_{k-1},M_k)$ is equivalent to conditioning on $(E_{k-1}, j_k)$. If $\sigma^2$ is known, the test is based on
\begin{equation}\label{eq:selModel_cond}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
\end{equation}
If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$.

We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail.

If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls. This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the null distribution we compare it against, which depends on $E_{k-1}, j_k,$ and $U_{k-1} = X_{E_{k-1}}'Y$.

Note the contrast between~\eqref{eq:selModel_cond} and~\eqref{eq:satModel}. In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on $n-1$-dimensional projection of $Y$. Conditioning on more information tends to sap the power of a test, and as we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests can pay a heavy price for this extra conditioning.

\subsubsection{Testing Against the Upper Model}

To avoid conditioning on the next model $M_k$, we could instead test $M_{k-1}$ against the alternative $M_{\infty}\setminus M_{k-1}$. In that case, we could modify~\eqref{eq:selModel_cond} and base the test on
\begin{equation}\label{eq:selModel_marg}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right).
\end{equation}
We could also replace $X_{j_k}'Y$ with any other test statistic $T(Y)$. As long as $T(Y)$ is measurable with respect to $(M_k, U_k)$, and the path algorithm is subpath-sufficient, the $p$-values are guaranteed to be independent on nulls. The next-$\lambda$ test for the lasso and the max-$t$ test in forward-stepwise regression with unknown $\sigma^2$ are both examples of this approach. Note that under resampling from the law in~\eqref{eq:selModel_marg}, the selected variable $j_k$ is random; by contrast, \eqref{eq:selModel_cond} conditions on $j_k$.

In the case of known $\sigma^2$, the greedy likelihood ratio statistic is the max-$z$ statistic: the maximum of $z_{j,E_{k-1}}$, the largest multivariate $z$-statistic of any held-out variable, controlling for the variables in the current model.

For the most part we do not pursue tests like those in Section~\ref{sec:identify}, which condition on the next model $M_k$. It is not necessary to condition on $M_k$ to obtain the FDR and FWER guarantees that we want to establish, and so we follow the rule of thumb that we should condition on as little information as possible. In Section~\ref{sec:sparseReg}, we empirically compare the max-$t$ test with the test that uses the same test statistic but also conditions on the identity of the next variable $j_k$. We call the resulting test the {\em max-$t$-identify test}. Its performance in our simulation is similar to that of the max-$t$ test.

\subsubsection{Classification of Proposed Tests}
Table \ref{tab:testsummary} classifies a number of different tests that have been proposed for least angle regression, lasso and forward stepwise regression. Some focus on the complete null hypothesis, others on the incremental null (as defined in Section \ref{sec:whichnull}).
Some use the saturated-model, and others the selected-model framework, as detailed in earlier in the section.
The saturated-model tests are simpler and do not require sampling, while the selected-model tests are generally more powerful. In the next section we compare the selected- and saturated-model settings in more detail.

\begin{table}
\begin{center}
\begin{tabular}{ | l | r|r| r}
        \hline
Procedure& Complete Null & Incremental Null\\
\hline
LAR/lasso& Covariance test/Saturated&Spacing test/Saturated \\
        & next-$\lambda$ test/Selected& \\
        \hline
Forward Stepwise &max-$t$, max-$t$-identify/Selected &Gaussian pivot/Saturated\\
\hline
\end{tabular}
\end{center}
\caption{\em Summary of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selected).
The max-$t$, max-$t$-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
pivot tests are given in \citet{taylor2014exact}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds.}
\label{tab:testsummary}
\end{table}
Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $N=50, p=10$ and two strong signals.
We plot the $p$-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
We see that the next-$\lambda$ test is the most powerful, especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
\begin{figure}
  \includegraphics[width=\textwidth]{figs/larexample.pdf}
  \caption[fig:larexample]{\em  Cumulative distribution functions of $p$-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
  \label{fig:larexample}
\end{figure}

\subsection{Comparison of Selected- and Saturated-Model Inference}\label{sec:bivariate}

We now briefly compare the computational, conceptual, and statistical advantages and disadvantages of selected- and saturated-model tests, focusing on the problem of sequential model selection. For a detailed general discussion, see \citet[][Section 5]{fithian2014optimal}.

Selected-model inference is typically more computationally involved than saturated-model inference. Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable. By contrast, selected-model tests typically require sampling from a truncated multivariate Gaussian distribution of dimension $p-|E_{k-1}|$.

For some applications, a major conceptual advantage of the saturated-model approach is that, when $M(E)$ is misspecified, $\theta^{E}$ is still a well-defined quantity on which we can perform exact inference. For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ could cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified. In the setting of sequential model selection, this advantage is less important: we are not looking for a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data. If there are many remaining strong signal variables to be found, then both $M_k$ and $M_{k-1}$ are badly misspecified; the more true this is, the stronger the grounds are for rejecting $M_{k-1}$.

\JTcomment{I would say our goal is a good stopping rule AND inference on parameters of selected (or perhaps saturated) model. We do not consider it here
because we don't know how to do it, not because it is not interesting.}

There are several major statistical benefits to selected-model inference. First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known. This is not an option for the saturated model because conditioning on both $\|Y\|^2$ and $\proj_{\eta_k}^\perp Y$ results in a degenerate conditional distribution for $\eta_k'Y$.

Second, we have seen in Section~\ref{sec:pValsIndep} that tests of the form~\eqref{eq:selModel_cond} or~\eqref{eq:selModel_marg} yield independent $p$-values under general conditions, allowing us to apply  the sequential stopping rules of~\citet{gsell2013sequential} and \citet{li2015accumulation}.

Finally, and perhaps most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we illustrate now with an extended example.

\begin{example}[Bivariate Regression with Identity Design]\label{ex:bivariate}
  Consider forward stepwise selection in a regression model with $n=p=2$, with known $\sigma^2=1$ and identity design 
\[
X = I_2=\begin{pmatrix} 1 & 0 \\ 0 & 1\end{pmatrix}.
\]

The forward stepwise path and the lasso path both select $j_1=1$ if and only if $|Y_1|>|Y_2|$. The selection event $A_1=\{|Y_1| > |Y_2|\}$ is shown in yellow in Figure~\ref{fig:bv_condSets}. If $j_1=1$ then
\[
M_0:\; Y\sim \cN(0,I_2), \quad \text{ and } \quad
M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, \; I_2\right).
\]

The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-$\lambda$ tests all coincide.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$. Thus, it must condition on $Y_2$ to eliminate the nuisance parameter $\mu_2$, comparing $Y_1$ to its null distribution given $A_1$ and the observed value of $Y_2$.

Figure~\ref{fig:bv_condSets} shows the conditioning sets for each model when $Y=(2.9, 2.5)$. Beside it, Figure~\ref{fig:bv_nullDists} shows the null conditional distribution of the test statistic $Y_1$ for each test. The $p$-values for the selected and saturated models are 0.007 and 0.3, respectively. These two plots are reproduced from \citet{fithian2014optimal}, in which the same example was presented in less detail.
\end{example}

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_condSets.pdf}
    \caption{The selected-model test conditions on $j_1=1$ (yellow region), while the saturated-model test also conditions on $Y_2=2.5$ to eliminate the nuisance variable $\mu_2$ (brown line segments).}
    \label{fig:bv_condSets}
  \end{subfigure}
  \hspace{.1\textwidth}
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_nullDists.pdf}
    \caption{Conditional distributions of $Y_1$ under
      $M_0:\; Y \sim \cN(0,I_2)$. The realized value $|Y_1|=2.9$ is
      quite large given that $j_1=1$. By
      contrast, $|Y_1|=2.9$ is not especially large once we 
      also condition on $Y_2=2.5$.}
  \label{fig:bv_nullDists}
  \end{subfigure}
  \caption{Contrast between the saturated-model and selected-model
    tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\L(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\L(Y_1  \gv Y_2, \; j_1=1)$. 
    When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.007 and 0.3, respectively.}
  \label{fig:bv}
\end{figure}

Figure~\ref{fig:bv} illustrates a general phenomenon in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak~\citep{lockhart2014significance}. Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=(4,4)$, a very strong signal. While the selected-model test has nearly perfect power, it is not uncommon for the saturated-model test to produce large $p$-values, even in the range of 0.5-0.9. These large $p$-values arise when there is a near tie between the variables.

\begin{figure}
  \centering
  % source code: bivariateSelVSat.R
  \includegraphics[width=.5\textwidth]{figs/bivariateSelVSat_rocCurve.pdf}
  \caption{Cumulative distribution function of $p_1(Y)$ for the selected- and saturated-model tests when $\mu=(4,4)$. Even though the signal is very strong, the saturated-model test results in large $p$-values in realizations like the one in Figure~\ref{fig:bv_nullDists}, where there is a near-tie between $|Y_1|$ and $|Y_2|$. By contrast, the selected-model test has nearly perfect power.}
  \label{fig:bv_rocCurve}
\end{figure}

Results in \citet{fithian2014optimal} show that the selected-model test is strictly more powerful when the selected model is correct; i.e., when $\mu_2=0$. Figure~\ref{fig:bv_powCurves} shows the power curve for each test when $\mu_2=0$ (left panel) and $\mu_2=4$ (right panel). While the selected-model test is more powerful when the selected model is correct, the difference between the two is relatively small. The difference is much more pronounced when $\mu_2=4$.

Note that if $\mu=(0,4)$, the incremental null is true, but the complete null is false. That is, the model $M_0 = M(\emptyset)$ is missing an important signal variable, but the missing variable is not $X_1 = (1,0)$. Because the saturated-model test is a valid test of the incremental null, its power is $\alpha=0.05$. By contrast, the selected-model test rejects about half of the time when $\mu=(0,4)$, detecting the fact that $M_0$ is a poor fit to the data.

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_0.pdf}
    %\caption{\WFcomment{Write caption here.}}
    %\label{fig:bv_powCurves_0}
  \end{subfigure}
  \hspace{.1\textwidth}
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_4.pdf}
    %\caption{\WFcomment{Write caption here.}}
  \end{subfigure}
  \caption{Power at level $\alpha=0.05$ for saturated- and selected-model tests at step 1, given that $j_1=1$. The power is plotted as a function of $\mu_1$, for two different values of $\mu_2$. When $\mu_2=0$, the selected-model test is strictly more powerful than the saturated-model test, but the difference is slight. By contrast, when $\mu_2=4$, the selected-model test is much more powerful. The dashed line shows $\alpha=0.05$.}
   \label{fig:bv_powCurves}
\end{figure}

Saturated-model $p$-values are generically non-independent. Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2, simulated under the global null $\mu=0$. Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-0.48$. By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Theorem~\ref{thm:suffCond}.

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Mon May  4 21:11:15 2015
\begin{table}[ht]
  \centering
  \begin{tabular}{l|ccccc|c}
    \multicolumn{7}{c}{Saturated-Model $p$-Values 
      (\% of $10^6$ Simulations)}\\[7pt]
    \hline
    \multicolumn{7}{c}{}\\[-1.5ex]
    \multicolumn{7}{c}{$p_2(Y)$}\\[5pt]
    ${\large p_1(Y)}$ & (0,0.2] & (0.2,0.4] & (0.4,0.6] & (0.6,0.8] & (0.8,1] & \textbf{Total} \\ 
    \hline
    (0,0.2] & 1.0 & 2.7 & 4.2 & 5.6 & 6.7 & 20.1 \\ 
    (0.2,0.4] & 1.4 & 3.4 & 4.5 & 5.2 & 5.5 & 20.0 \\ 
    (0.4,0.6] & 2.3 & 4.3 & 4.7 & 4.5 & 4.2 & 20.0 \\ 
    (0.6,0.8] & 4.2 & 5.4 & 4.3 & 3.4 & 2.7 & 20.0 \\ 
    (0.8,1] & 11.1 & 4.3 & 2.3 & 1.4 & 1.0 & 20.0 \\ 
    \hline
    \textbf{Total} & 19.9 & 20.0 & 20.0 & 20.1 & 20.0 & 100.0 \\ 
    \hline
  \end{tabular}
  \caption{Two-way contingency table of saturated-model $p$-values $(p_1(Y), p_2(Y))$ for Example~\ref{ex:bivariate}, after binning into cells of height and width 0.2. We report the percentage of $p$-value pairs falling into each cell out of one million simulations from the global null hypothesis, $\mu=0$. Both $p$-values are marginally uniform but strongly dependent, with a correlation of $-0.48$.}
\label{tab:bv_twoWayTable}
\end{table}

\section{Computation}
\label{sec:computation}

Until this point, we have deferred discussing the question of how to actually compute the rejection cutoffs for the max-$t$ test, next-$\lambda$ test, or any of the other tests we have discussed. In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law
\begin{equation}\label{eq:genDist}
\L(T_k(Y) \,\mid\, M_{k-1}(Y), U_{k-1}(Y)),
\end{equation}
under the null hypothesis that $F\in M_{k-1}$. Because $U_{k-1}$ is sufficient for $Y$ under the null, this law is completely known. Thus, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from its null distribution given $M_{k-1}(Y), U_{k-1}(Y)$, we can numerically approximate the $p$-value to whatever precision we desire.

In many cases it is easy to sample $Y$ given $U_{k-1}$, so computational difficulties only arise insofar as the selection event $\{M_{k-1}(Y)=M\}$ is difficult to condition on. Due to the generality in which we have posed the problem, we cannot give a general prescription for how to carry out this sampling. However, we give some details for the important case of linear regression in Appendix~\ref{sec:linRegComputation}.

For a completely general selection algorithm, we might expect severe computational difficulties when we attempt to condition on the null model at stage $k$, $M_{k-1}(Y)$. For example, in forward stepwise regression, $M_{k-1}$ is the accumulated result of $k-1$ steps of greedy variable selection. At each step $s<k$, the selection of variable $j_s$ is a result of comparing its $t$-statistic with the $t$-statistics of the other $p-k$ inactive variables. Based on this logic, we might expect to accumulate at least $O(p)$ constraints at each step, resulting in $O(kp)$ constraints after $k$ steps.

For lasso linear regression, the situation seems even more dire: for any fixed value of $\lambda$, the event $\{\supp(\hat\beta^\lambda) = E\}$ is a union of up to $2^{|E|}$ polytopes in $\R^n$, each satisfying $O(p)$ constraints \citep{lee2013exact}. \footnote{To finesse this issue, \citet{lee2013exact} propose conditioning on the signs of $\hat\beta^\lambda$ as well. If we do this, the conditioning set becomes a single polytope.}. But the event $\{E_{k-1}(Y)=E\}$ depends on $\supp(\hat\beta^\lambda)$ for every $\lambda > \lambda_k$ --- i.e., on the solution to {\em infinitely many} lasso problems.

Remarkably, however, in both forward-stepwise and lasso linear regression, the computational problem actually becomes easier when we move to the sequential setting considered here. Indeed, Section~\ref{sec:constraints} shows that for both forward-stepwise regression and lasso linear regression --- and their respective generalizations to exponential family models --- conditioning on the event $\{E_{k-1}=E\}$, for any $k$, only introduces $2p$ linear constraints, taking the form of a rectangular constraint on the sufficient statistics for the inactive variables.

\subsection{Constraints}\label{sec:constraints}
Happily, it turns out that the conditioning polytope can be expressed with only $2(p-k)$ constraints in forward stepwise regression and the lasso, as we show here. This increases the speed of both accept/reject and hit-and-run sampling for our problem.

We begin by proving the result for forward stepwise linear regression, followed by a more general result for the forward stepwise likelihood path in exponential families.

\begin{theorem}
  Assume the model path $M_{0:d}$ is obtained by forward stepwise 
  linear regression. Then, for a candidate active set $E$ of size $k$, 
  the set $A = \{E_k = E, \;X_E'Y = u\}$ is characterized 
  exactly by the constraints $X_E'Y=u$ and
  \[
  v_j^-(E,u) < X_j'Y < v_j^+(E,u), \quad\forall j \notin E,
  \]
  for $v_j^-$ and $v_j^+$ given explicitly in
  Equations~(\ref{eq:vMinus_FS}--\ref{eq:vPlus_FS}).
  Thus, $A$ corresponds exactly to 
  a set with $2(p-k)$ linear inequality constraints and $k$
  linear equality constraints on $Y$.
\end{theorem}

\begin{proof}
  At stage $i < k$, the next variable to enter is the one with maximal correlation with the residual vector; i.e.
  \[
  j_{i+1} = \argmax_{j \notin E_i} 
  \frac{\left| X_j ' (Y - X_{E_i}\hat\beta^i) \right|}
  {\|\proj_{E_i}^\perp X_j\|}
  \]
  Because forward stepwise regression satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on the $k$th active set $E_k$ and its sufficient statistics $X_{E_k}'Y$. On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\beta^i$, and $\proj_{E_i}^\perp$ depend only on $X_E'Y$. For brevity, write 
  \[
  C_i^* = \max_{j \notin E_i} 
  \frac{\left| X_j ' (Y - X_{E_i}\hat\beta^i) \right|}
  {\|\proj_{E_i}^\perp X_j\|}.
  \]
  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th variable added.

  If $X_E'Y$ is fixed at $u$, and $j \notin E$, then the condition for 
  $X_j$ {\em not} to enter at step $i+1 < k$ is
  \[
  \frac{\left| X_j' (Y - X_{E_i}\hat\beta^i) \right|}
  {\|\proj_{E_i}^\perp X_j\|} 
  \leq C_i^*,
  \]
  or equivalently,
  \begin{equation}\label{eq:noEnterBounds_FS}
    X_j' X_{E_i}\hat\beta^i -
    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|
    \;\;\;\leq\;\;\;
    X_j'Y
    \;\;\;\leq\;\;\;
    X_j' X_{E_i}\hat\beta^i +
    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|, 
  \end{equation}
  so the set $A$ is equivalent to 
  Equation~\eqref{eq:noEnterBounds_FS} holding
  for every $i < k$ and $j \notin E$. Nominally, this gives $2k(p-k)$
  linear inequality constraints to satisfy, but most of them are
  non-binding. On $A$, the upper and lower bounds
  in~\eqref{eq:noEnterBounds_FS}
  are all known functions of $E$ and $u$, so we can set
  \begin{align}\label{eq:vMinus_FS}
    v_j^-(E,u) &= \max_{0 \leq i < k} \;\;X_j' X_{E_i}\hat\beta^i -
    C_i^*\|\proj_{E_{i}}^\perp X_{j}\| \\
    \label{eq:vPlus_FS}
    v_j^+(E,u) &= \min_{0 \leq i < k} \;\;X_j' X_{E_i}\hat\beta^i +
    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|.
  \end{align}  
\end{proof}

As we see below, the same result holds for the much larger class of forward stepwise likelihood paths in exponential family models.

\begin{theorem}
  Assume that $M_\infty$ is an exponential family model
  of the form
  \[
  Y \sim \exp\{ \theta'U(y) - \psi(\theta) \}\,d\nu(y),
  \]
  with $\Theta \sub \R^p$ convex, and assume that the model path $M_{0:d}$ is given by the forward stepwise likelihood path with
\begin{equation}
j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
E_k = E_{k-1} \cup \{j_k\}.
\end{equation}

  Then, for a candidate active set $E$ of size $k$, the set $A = \{E_k = E, \;U_E = u\}$ is characterized exactly by the constraints $U_E = u$ and
  \[
  v_j^-(E,u) \leq U_j(Y) \leq v_j^+(E,u), \quad\forall j \notin E,
  \]
  for $v_j^-$ and $v_j^+$ given in Equations~(\ref{eq:vMinus_FSL}--\ref{eq:vPlus_FSL}).
  Thus, $A$ corresponds exactly to a set with $2(p-k)$ linear inequality constraints and $k$ linear equality constraints on $U(Y)$.
\end{theorem}

\begin{proof}
  For a set $E\sub \{1,\ldots,p\}$ define
  \[
  \ell_E^*(U_E(Y)) 
  = \sup\left\{\theta'U(Y) - \psi(\theta) :\; \theta\in \Theta(E) \right\}
  = \sup\left\{\theta_E'U_E(Y) - \psi_E(\theta_E):\; \theta\in\Theta(E) \right\},
  \]
  where $\psi_E:\; \R^{|E|} \to \R$ is defined such that $\psi(\theta)=\psi_E(\theta_E)$ if $\theta_{\setminus E}=0$. Note that $\ell_E^*$ is the convex dual of the convex function $\psi_E$, so $\ell_E^*$ is also convex. \JTcomment{It would be convex even if $\psi_E$ were not convex. No? I also think $U$ in the definition of $\ell^*_E$ can be just $U(Y)$.} At stage $i<k$, the next variable to enter is the one that maximizes $\ell_{E_i\cup \{j\}}^*(U(Y))$.

  Because the forward stepwise likelihood path satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on $E_k$ and its sufficient statistics $U_{E_k}(Y)$. On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\theta^i$, and $\proj_{E_i}^\perp$ depend only on $U_E(Y)$. For brevity, write 
  \[
  C_i^* = \max_{j \notin E_i} 
    \sup \left\{ \theta ' U - \psi(\theta):\; \theta \in \Theta(E_i\cup \{j\}) \right\}.
  \]
  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th variable added.

  If $U_E(Y)$ is fixed at $u$, and $j \notin E$, then the condition for 
  variable $j$ {\em not} to enter at step $i+1 < k$ is
  \[
  \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*.
  \]
  Note that $\ell_{E_i\cup \{j\}}^*$ depends only on the sufficient statistics $U_{E_i}(Y)$ --- which is fixed, on $A$ --- and the remaining sufficient statistic $U_j(Y)$. Since $\ell_{E_i\cup\{j\}}^*$ is convex, the set
  \begin{equation}\label{eq:noEnterBounds_FSL}
  \{U_E=u, \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*\} 
  = \{ U_E=u, v_{j,i}^-(E,u) \leq U_j \leq v_{j,i}^+(E,u)\},
  \end{equation}
  where $v_j^{-}$ and $v_j^+$ are the two values where $\ell_{E_i\cup \{j\}}^*(U)=C_i^*$.

  The set $A$ is equivalent to 
  Equation~\eqref{eq:noEnterBounds_FSL} holding
  for every $i < k$ and $j \notin E$. Nominally, this gives $2k(p-k)$
  linear inequality constraints to satisfy, but most of them are
  non-binding. On $A$, the upper and lower bounds
  in~\eqref{eq:noEnterBounds_FSL}
  are all known functions of $E$ and $u$, so we can set
  \begin{align}\label{eq:vMinus_FSL}
    v_j^-(E,u) &= \max_{0 \leq i < k} \;\;v_{j,i}^-(E,u)\\
    \label{eq:vPlus_FSL}
    v_j^+(E,u) &= \min_{0 \leq i < k} \;\;v_{j,i}^-(E,u)
  \end{align}
\end{proof}

A similar result also holds for the class of $\ell_1$-regularized exponential family models including the graphical LASSO.

\begin{theorem}
  Assume that $M_\infty$ is an exponential family model
  of the form
  \[
  Y \sim \exp\{ \theta'U(y) - \psi(\theta) \}\,d\nu(y),
  \]
  with $\Theta \sub \R^p$ convex, and assume
  that the model path $M_{0:d}$ is given 
  by the ever-active set for the $\ell_1$-penalized problem
  \begin{equation}\label{eq:regProblem}
  \hat\theta^\lambda = \argmin_{\theta\in \Theta} 
  -\ell(\theta) + \lambda\|\theta\|_1,
  \end{equation}
  for $\lambda\in \Lambda \sub [0,\infty)$.

  Then, for a candidate active set $E$ of size $k$, 
  the set $A = \{E_k = E, \;U_E = u\}$ is characterized 
  exactly by the constraints $U_E = u$ and
  \[
  v_j^-(E,u) \leq U_j(Y) \leq v_j^+(E,u), \quad\forall j \notin E,
  \]
  for $v_j^-$ and $v_j^+$ given explicitly in
  Equations~(\ref{eq:vMinus_L1}--\ref{eq:vPlus_L1}).
  Thus, $A$ corresponds exactly to 
  a set with $2(p-k)$ linear inequality constraints and $k$
  linear equality constraints on $U(Y)$.
\end{theorem}

\JTcomment{Is there a reason you have $\psi$ and $\ell$ above instead of just $\ell$? I think the $(E,t)$ above should be $(E,u)$. It seems you never come back to $\psi$.}
\WFcomment{Fixed t/u. $\psi$ is the log-proportionality constant, I think we need it. Maybe I'm misunderstanding your question?}
\JTcomment{I guess I meant that I am assuming that
$$
\ell(\theta) = \theta'U(y) - \psi(\theta)
$$
but I don't see how I would conclude that from what is written.
In principle, as written the $\psi$ above might be some other function but
(44) and (45) imply that the objective function is of the form
$$
\ell(\theta) = \tilde{\psi}(\theta) - \theta'U(y)
$$
for some $\tilde{\psi}$. In the proof when you say $\mathbb{E}_{\hat{\theta}^{\lambda}}[U_j]$ I am assuming this is $e_j^T\nabla \psi (\hat{\theta}^{\lambda})$ soshould I conclude that $\tilde{\psi}=\psi$?
}

\begin{proof}
  Note that, because exponential family likelihoods are concave in their natural parameters, the problem in~\eqref{eq:regProblem} is convex. Again, once we know the sufficient statistics $U_E(Y)$ for model $k$, and that $E_k=E$, we know that
  \[
  \hat\theta^\lambda = \hat\theta^{(E,\lambda)}
  \]
  for every $\lambda \geq \lambda_k$, so we know the entire path of fits up to and including $\lambda_k$. But then, excluding variable $j \notin E$ at Lagrange parameter $\lambda$ is equivalent to
  \[
  \lambda \geq 
  \left| \pardd{\ell(\hat\theta^\lambda)}{\theta_j} \right|
  = \left|U_j - \E_{\hat\theta^\lambda}[U_j]\right|
  \]
  On $A$, for $\lambda \geq\lambda_k$, 
  $\E_{\hat\theta^\lambda}[U_j]$ is a known function of $E$ and $u$,
  so we can set
  \begin{align}\label{eq:vMinus_L1}
    v_j^-(E,u) &= \sup_{\lambda \geq \lambda_k} \;\; 
    \E_{\hat\theta^\lambda}[U_j] - \lambda \\
    \label{eq:vPlus_L1}
    v_j^+(E,u) &= \inf_{\lambda \geq \lambda_k} \;\;
    \E_{\hat\theta^\lambda}[U_j] + \lambda .
  \end{align}
\end{proof}

\section{Simulation: Sparse Linear Regression}
\label{sec:sparseReg}

Next we compare several model selection procedures in simulation. We simulate from a linear regression model with $n=100$ observations and $p=40$ variables. The design matrix $X\in\R^{n\times p}$ is a random Gaussian design with pairwise correlations of 0.3 between predictor variables.

The columns of $X$ are normalized to have length 1, and we simulate from $Y \sim \cN(X\beta,I_n)$, using a seven-sparse model with signal-to-noise ratio 5:
\[
\beta_j = \left\{\begin{matrix}5 & j = 1,\ldots,7\\ 0 &
    j>7\end{matrix}\right.
\]
We use known $\sigma^2=1$, so that we can compare the saturated-model test with the selected-model test. For our selection algorithm, we use the entire forward-stepwise path, for all 40 steps.

\subsection{Single-Step $p$-Values}

For each step we compute one-step selected-model and saturated-model $p$-values, as well as nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable. Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first ten steps, conditional on the event that the null hypothesis is false. It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values. The nominal $p$-values are also quite powerful, but they do not have the correct level.

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
  \caption{\em  CDFs of nominal (black), MaxZ (red), MaxT (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$. The MaxT approaches are much more powerful than the saturated-model test. The nominal test appears to be powerful, 
  but is not $U(0,1)$ under the null, as shown below in Figure \ref{fig:simulation_null_true}.}
  \label{fig:simulation_null_false}
\end{figure}

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
  \caption{\em  CDFs of nominal (black), MaxZ (red), MaxT (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.  The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
  \label{fig:simulation_null_true}
\end{figure}

\begin{figure}[h]
  \centering
  % source code: ??? for simulation, sparseSim.R for plotting
  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
  \caption{\em CDFs of nominal (black), MaxZ (red), MaxT $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model. Here, none of the methods produce uniform $p$-values. The null hypothesis is false in most cases and so --- in our model-centric point of view --- rejection is the desired outcome.}
  \label{fig:simulation_noise_var}
\end{figure}

Figure~\ref{fig:simulation_null_true} shows the distribution of $p_k$ for $k = 8, \ldots, 17$, given that the null hypothesis tested at step $k$ is correct (i.e., that $k_0< k$). Because the correct model is seven-sparse, $k=8$ is the first index for which the null can possibly be true. All but the nominal $p$-values are uniform by construction, but the nominal $p$-values are highly anti-conservative, as expected.

Finally, as a warning against misinterpretation of our method, we include Figure~\ref{fig:simulation_noise_var} showing the first ten $p$-values for each method, conditional on event that the variable added at step $k$ is a noise variable in the full model. Now, none of the $p$-values are uniform.

This is not an error in our implementation of the method, but rather a consequence of our ``model-centric'' point of view. If we try to add a noise variable to the model before we have included all the signal variables, then we are testing a false null hypothesis. The test rejects because there is much more signal to find, and as such, it is entirely appropriate for us to reject the null and continue growing the model.

\subsection{Model-Selection Performance}

If we combine the saturated-model or selected-model $p$-values with one of our three stopping rules, we can evaluate the model-selection performance of each method in terms of:
\begin{itemize}
\item its probability of selecting a correct model or $p_{\text{screen}}$,
\item its model-wise FWER,
\item its model-wise FDR, and
\item its variable-wise FDR, where we use $V^{\text{full}}= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of $V=(\hk-k_0)_+$.
\end{itemize} 
The last measure of performance is not explicitly controlled by any of the selective-inference methods, but we might nevertheless hope to perform reasonably.

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Fri May  1 09:29:28 2015
%\begin{table}[ht]
 % \centering
 % \begin{tabular}{llcccc}
  %  \hline
  %  Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ an
 %   & $\text{FDR}_{\text{mod}}$ 
  %  & $\text{FDR}_{\text{var}}$ \\ 
 %   \hline
    %Selected & Basic & .290 & .000 & .002 & .039 \\ 
   % Selected & Forward & .559 & .027 & .020 & .066 \\ 
   % Selected & Strong & .041 & .014 & .008 & .042 \\ 
  %  \hline
   % Saturated & Basic & .000 & .000 & .000 & .028 \\ 
  %  Saturated & Forward & .014 & .000 & .000 & .030 \\ 
  %  Saturated & Strong & .000 & .000 & .000 & .032 \\ 
  %  \hline
  %  Knockoffs & & .000 & --- & --- & .231 \\ 
  %  \hline
 % \end{tabular}
 % \caption{\WFcomment{Comment on results once they are correct. These numbers are wrong right now.}}
%\end{table}

\begin{table}[ht]
\centering
\input{tables/error_rates_05.tex}
\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}. Results with theoretical guarantees are in \guarantee{bold} type.}
\label{tab:stopping05}
\end{table}

\begin{table}[ht]
\centering
\input{tables/error_rates_20.tex}
\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}. Results with theoretical guarantees are in \guarantee{bold} type.}
\label{tab:stopping20}
\end{table}

We see that basicStop and forwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, 2) the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism. Although not guaranteed by the theory, the $p$-values from the saturated model do control FWER and model-wise FDR in this example, 4) the max-$t$ rules show higher probabilities of selecting the true model, especially using forwardStop, 5) the knockoff method does not yield variable-wise FDR control.

\section{Futher Examples and Extensions}\label{sec:further-examples}

Although we have focused our discussion on the case of linear regression --- a very common and important application of sequential model selection --- our theory is quite general and applies to many other parametric and nonparametric models. To give a sense of the broader applicability, we now discuss one parametric and one nonparametric example, each of which fit nicely into our theory.

\subsection{Decision Trees}
In this section we discuss another application of adaptive sequential inference, to binary decision trees. Specifically, the classification and regression tree  (CART)  method builds a decision tree in a top down manner, finding the best available split at each node.  When considering splitting two daughter nodes, the generic CART algorithm splits the nodes in an arbitrary order, only stopping when a node reaches a minimum size.  We instead consider a  special version of CART  algorithm:  the "best first"  method, which orders splits by their achieved reduction in loss. This procedure is used, for example in the R package for gradient boosting, called {\tt gbm}. The best-first algorithm is thus a sequential selection  procedure, to which we can apply the results in this paper. Hence we can obtain independent, sequential $p$-values for each successive split in the decision tree.

Given predictor matrix $X \in \R^{n\times p}$ and binary response vector $Y\in \{0,1\}^n$, we begin with the intercept-only model $\logit \P(Y_i \mid X_i = x) \equiv \beta_0$.

At the first step we choose a splitting variable $j_1$ and split point $t_1$ which is a midpoint between two observed values of variable $j$, leading to a model of the form
\begin{equation}
\logit \P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\}.
\end{equation}
While there are various ways to choose among potential variables and split points, we assume for the sake of simplicity that $(j_1,t_1)$ are chosen to maximize binomial likelihood (or equivalently, to minimize binomial deviance).

At the second stage, we then split one of the two daughter nodes, choosing the split that gives the greatest increase in likelihood. If (say) we split the first node using predictor $j_2$ at split point $t_2$, it leads to the interaction model of the form
\begin{equation}
\logit \P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\} + \beta_2 1\{x_{j_1} \leq t_1, x_{j_2} > t_2\}
\label{eqn:intmodel}
\end{equation}

After $k$ steps we obtain a logistic regression model with $k$ features, each of which is an interaction of step functions. The model $M_k$ is thus an exponential family whose complete sufficient statistics are the sums of $Y_i$ in each leaf in the tree.

The best-first search for split points can be expressed as a modified form of greedy forward stepwise logistic regression in which the set of candidate features changes at each step, depending on what features have already been included. Nevertheless, the proof of Proposition~\ref{prop:forwardSSP} goes through essentially without modification to show that this algorithm satisfies the SSP, and the greedy likelihood ratio test gives $p$-values that are independent on nulls\footnote{Since the test statistic is discrete, we need to randomize if we want truly exact $p$-values}.

This development suggests a possible modification to the CART algorithm in which the tree is ``pruned'' using a rule like ForwardStop applied. It could be interesting to investigate whether cross-validation using the target FDR $\alpha$ as a tuning parameter might yield better results in predictive performance, compared to (say) using the tree depth or minimum leaf size as a tuning parameter.

\subsection{Nonparametric Changepoint Detection}

Consider a problem in which we observe a time series $(Y_1,\ldots,Y_T)\in \cY^T$, and wish to detect times $t_1,\ldots,t_k$ where the distribution of $Y_t$ changes. For example, given a historical document, we may wish to investigate whether the entire document was written by the same author or different sections were written by different authors, by quantitatively comparing the writing style across different sections. For example, $Y_t$ could could represent the usage distribution in section $t$ over ``context-free'' words such as articles, conjunctions, and prepositions \citep[][see e.g. ]{chen2015graph}.

For simplicity we assume independence between values of $t$ and model the distribution as piecewise constant between $k$ unknown changepoints. For an active set of $k$ changepoints $E\sub \{1,\ldots, T-1\}$, let $t(j,E)$ denote the $j$th smallest element of $E$, and let $t(0,E)=0$ and $t(k+1, E)=T$. Then, the nonparametric model with changepoints $E$ is given by
\begin{equation}\label{eq:nonpar-model}
  M(E):\; Y_t \simind F_j, \quad \text{ for } \;\;t(j,E) < t \leq t(j+1,E).
\end{equation}

For $t_2>t_1$ let $V(Y;\, t_1, t_2)\in \cY^{t_2-t_1}$ denote the order statistics of $(Y_{t_1+1},\ldots, Y_{t_2})$. Then the complete sufficient statistic for $M(E)$ is
\[
U(Y; \, E) = 
\bigg(
V(Y; t(0, E), t(1, E)), \ldots, V(Y;\, t(k, E), t(k+1,E))
\bigg).
\]
Under $M(E)$, resampling from $Y$ given $U$ amounts to randomly permuting $(Y_{t(j,E)+1}, \ldots, Y_{t(j+1,E)})$ independently for each $j=0,\ldots,k$.

For $s \notin E$, testing $M(E)$ against $M(E \cup \{s\})$ amounts to a two-sample test comparing the observations coming immediately before $s$ against the observations coming immediately after $s$. Specifically, if $t(j, E) < s < t(j+1, E)$, let $W(Y; \, E, s)$ denote any two-sample test statistic that is measurable with respect to the samples $V(Y; t(j, E), s)$ and $V(Y; s, t(j+1, E))$, and large when there is strong evidence that the two samples come from different distributions.

We can build a model in a greedy fashion beginning with $E_0 = \emptyset$ and then at step $k=1,\ldots,d$ setting
\begin{equation}
  t_k = \argmax_t \;\; W(Y; \, E_{k-1}, t), 
  \quad \text{ and } \quad E_k = E_{k-1} \cup \{t_k\}.
\end{equation}
This path algorithm satisfies the SSP because $W(Y; \, E_{i}, t_j)$ is measurable with respect to $U_k=U(Y; \, E_k)$ for every $i<j\leq k$. Thus, by inspecting $U_k$ we can determine in what order the first $k$ changepoints were added.

By analogy to the greedy likelihood ratio statistic, a natural choice of test statistic at step $k$ is
\[
W_k^*(Y) = \max_t W(Y; \, E_{k-1}, t),
\]
which is measurable with respect to $\sF(E_k, U_k)$ by construction. Thus, if we base the step-$k$ $p$-value $p_k$ on the law
\begin{equation}\label{eq:condPerm}
\L\left(W_k^* \mid E_{k-1}, U_{k-1}\right),
\end{equation}
randomizing to obtain an exactly uniform $p$-value under $M(E_{k-1})$, we will obtain an exact $\sF_k$-measurable $p$-value. As a result, the $p$-values will be independent on nulls per Theorem~\ref{thm:suffCond}.

Note that sampling from the law in~\eqref{eq:condPerm} can be carried out by permuting subsequences $(Y_{t(j,E)+1}, \ldots, Y_{t(j+1,E)})$ and accepting the permutation only if it gives the same $E_{k-1}$ as we get in the original data. When this simple accept/reject algorithm is impractical we may need to resort to an MCMC strategy. We do not pursue these computational matters here.

\section{Discussion}
\label{sec:discussion}

It is a commonplace that ``essentially all models are wrong, but some are useful'' \citep{box1987empirical}. In essence, a statistical model is useful if it is large enough to capture the most important features of the data, but still small enough that inference procedures can achieve adequate power and precision. Apart from theoretical considerations, the only way to know whether a model is large enough is to test it using available data.

Although model-checking is commonly recommended to practitioners as an important step in data analysis, it formally invalidates any inferences that are performed with respect to the model selected. Our work takes a step in the direction of reconciling that contradiction, but there are important questions left to be resolved. In particular: which sorts of model misspecification pose the greatest threat to our inferential conclusions, and how powerful are our tests against these troublesome sources of misspecification?

In future work we also plan to address the matter of performing selective inference for parameters of the model actually selected. In principle, we already know how to do this using the framework of \citet{fithian2014optimal} --- we simply condition on the event that $M_{\hk}$ is selected and perform inference with respect to $M_{\hk}$ --- but the task may be computationally awkward in general.

Open-source code is available at this article's github repository at:
\begin{center}
\texttt{www.github.com/wfithian/sequential-selection}
\end{center}

\section*{Acknowledgments}

The authors are grateful for stimulating and informative conversations with Stefan Wager, Lucas Janson, Trevor Hastie, Yoav Benjamini, and Larry Brown. \WFcomment{Funding}.

\bibliographystyle{plainnat}
\bibliography{biblio}

\newpage

\begin{appendix}

\section{Proof of Proposition~\ref{prop:counterexample}}

In this section we prove Proposition~\ref{prop:counterexample}, restated below for convenience.

\counterex*

\begin{proof}
Denote the events where we choose the two possible orderings as $A_{123}$ and $A_{213}$, respectively. Because the variables are orthogonal, the max-$z$ statistic for any model $M(E)$ is 
\[
|z_E^*|=\max_{j\notin E} |Y_j|.
\]

At each step we use the selective max-$z$ test conditioning on the current null model, where $p_k$ is small when $|z_{E_{k-1}}^*|$ is large compared to the law
\[
\L(|z_{E_{k-1}}^*| \,\mid\, E_{k-1}, Y_{E_{k-1}}).
\]

Note that whichever of $A_{123}$ and $A_{213}$ occurs, $E_2=\{1,2\}$ and $|z_{E_2}^*|=|Y_3|$. Thus, $E_2$ is nonrandom and $p_3=2\Phi(-|Y_3|)$, the usual two-sided normal $p$-value based on $Y_3$. Finally, note that $k_0=2$ on $A_{123}$ and $k_0=1$ on $A_{213}$.

We have
\begin{align*}
  FWER 
  &= \P(A_{123})\P(p_1<\alpha, p_2<\alpha, p_3<\alpha \mid A_{123})
  + \P(A_{213})\P(p_1<\alpha, p_2<\alpha \mid A_{213})\\
  &= \alpha \P(p_1<\alpha, p_2<\alpha \mid A_{123})
  + (1-\alpha) \P(p_1<\alpha, p_2<\alpha \mid A_{213}),
\end{align*}
using the fact that $A_{123}$ corresponds exactly to the event that $p_3<\alpha$. As $K\to\infty$, we have $z_\emptyset^*=z_{\{1\}}^*=Y_2>K/2$ with probability tending to 1. Thus, $p_1,p_2\approx 0$ on $A_{123}$, while $p_1\approx 0$ on $A_{213}$. Continuing the calculation, then,
\begin{align*}
  \lim_{K\to\infty} FWER 
  &= \alpha\cdot 1 + (1-\alpha)\cdot \P(p_2 < \alpha \mid A_{213})\\
  &= \alpha + (1-\alpha)\alpha = 2\alpha-\alpha^2.
\end{align*}
\end{proof}

\section{Computational Details for Linear Regression}\label{sec:linRegComputation}

As discussed in Section~\ref{sec:singleStep}, we will set
\[
p_k(Y) = p_{k, E_{k-1}(Y)}(Y),
\]
where $p_{k,E}(Y)$ for each $E\sub \{1,\ldots, p\}$ is a conditional $p$-value for the {\em fixed} null model $M(E):\, Y \sim N(X_E\beta, \sigma^2)$, valid on the selection event $A=\{E_{k-1}(Y)=E\}$.

If $\sigma^2$ is known, then $U_{k-1}(Y) = X_{E}'Y$ on $A$, so our task is to sample under the null from
\begin{equation}\label{eq:truncNorm}
\L(Y \mid X_E'Y, Y\in A),
\end{equation}
which is a multivariate Gaussian distribution supported on the hyperplane of dimension $n-|E|$ with $X_E'Y$ fixed at its realized value, and truncated to the event $A$.

If $\sigma^2$ is unknown, then $U_{k-1}=(X_E'Y,\, \|Y\|^2)$ on $A$ and we must sample under the null from
\begin{equation}\label{eq:truncUnif}
\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A),
\end{equation}
which is a uniform distribution supported on a sphere of dimension $n-|E|-1$ defined by fixing $X_E'Y$ and $\|Y\|^2$ at their realized values, and truncated to the event $A$. For the simulations in this article, we use a hybrid of two strategies:

\paragraph{Accept/Reject Sampler} In either of these two cases, sampling would be very easy if $A$ were the entire sample space $\R^n$; the only reason we might encounter difficulty is that $A$ could be an irregular set that is difficult to condition on. However, we can always sample from $\L(Y \mid U)$ and then keep each new sample $Y$ if it lies in the selection event.

\paragraph{Hit-and-Run Sampler} \citet{fithian2014optimal} propose sampling from laws like that in~\eqref{eq:truncNorm} via a hit-and-run sampler for the $n-|E|$-dimensional multivariate Gaussian distribution: at each step, to generate $Y^{t+1}$ from $Y^t$, we choose a uniformly random unit vector $\nu$ in the hyperplane, and resample $\nu'Y$ conditional on the constraints $\proj_{\nu}^\perp Y^{t+1}=\proj_{\nu}^\perp Y^t$ and the constraint $Y^{t+1}\in A$. If $A$ is a polytope this amounts to resampling a univariate Gaussian distribution with known mean and variance, truncated to an interval defined by maximizing and minimizing $\nu'Y^{t+1}$ subject to linear constraints. For the case with $\sigma^2$ unknown, the sampler is a bit more involved; see \citet{fithian2014optimal}.

Figure \ref{fig:comparison} compares the two samplers in an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
The ``exact'' answer was determined from bootstrap sampling with $20,000$ replications.
The Figures show the bias, variance and mean squared error (MSE) of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
\begin{figure}[htp]
\centering
  \includegraphics[width=\textwidth]{figs/comparison.pdf}
  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.  Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.
}
\label{fig:comparison}
\end{figure}
We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for 8000 samples. As we move further along the path, the set $A$ becomes more and more constraining, so that we must discard the vast majority of the samples we generate. Thus we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low for accept/reject to be practical.

\JTcomment{I don't really see the point of Figure 5. It is abundandtly clear
that if feasible an IID sample is preferable over an MCMC sample. It is
just as easy to make plots where accept/reject will take forever to get even one sample point.}

In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal. When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach. We exploit the fact that each move along a random direction  has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedron. Unlike accept/reject sampling, which produces independent samples,  the hit and run method produces dependent samples. Hence we must run it for longer with an initial burn in period, and hope that it has mixed sufficiently well.

The task is made much simpler by the fact that the selection event can we written as a polytope $A=\{y: \Gamma y \geq u\}$. Hence we don't need to run the entire forward stepwise procedure on $Y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check if $\Gamma Y^* \geq u$. In the case of forward stepwise and lasso regression, $\Gamma$ has $2(p-k)$ rows after $k$ steps, as we see next.

\end{appendix}

\end{document}